### 2024-04-23 assignments/assignment-2.tex
assignments-project

\documentclass[12 pt]{article}

% Packages for mathematical symbols and equations
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{color}
\usepackage{setspace}
\usepackage{acro}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{../../../tex-packages/physics}
\usepackage{../../../tex-packages/qcircuit}
\usepackage[framemethod=tikz]{mdframed}
\usepackage{cancel}
\usepackage{empheq}
\usepackage{tikz}
\usepackage{color}
\usepackage[english]{babel}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{adjustbox}
\usepackage{float}
%\numberwithin{equation}{section}
\usepackage{bm}
\usepackage{wrapfig}
\usepackage{relsize}
\usepackage[letterpaper ,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=1.75cm]{geometry}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usetikzlibrary{calc,trees,positioning,arrows,chains,shapes.geometric,%
    decorations.pathreplacing,decorations.pathmorphing,shapes,%
    matrix,shapes.symbols}
\include{acronyms}
%\geometry{top=1.3in,bottom=1.3in}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
  urlcolor=black}

% Set page margins

\title{Quantum Computation and Quantum Information\\
Assignment 2}
\author{Gabriel Balarezo}
\date{\today}

\begin{document}

\maketitle

\begin{mdframed}[backgroundcolor = gray!30,
  frametitle = Exercise 1.2]
  Explain how a device which, upon input of one of two non-orthogonal quantum states $\ket{\psi}$ and $\ket{\varphi}$ correctly identified the state, could be used to build a device which cloned the states $\ket{\psi}$ and $\ket{\varphi}$, in violation of the non-cloning theorem. Conversely, explain how a device for cloning could be used to distinguish non-orthogonal quantum states.
\end{mdframed}

Given a device that can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ (without measurement), we can then design a quantum circuit
that clones the states $\ket{\psi}$ and $\ket{\varphi}$ as we like.\\ 
First, let's define the non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ as follows:
\begin{align}
  \ket{\psi} & \in \{\ket{\psi_+}, \ket{\psi_-}\} \\
  \ket{\varphi} & \in \{\ket{\varphi_+}, \ket{\varphi_-}\}
\end{align}
then, there exists a single-qubit operator, i.e. a rotation such that:

\begin{align*}
  U(\phi, \theta) \ket{\psi_+} = \ket{0} \\
  U(\phi, \theta)\ket{\psi_-} = \ket{1} \\
  U(\phi, \theta)\ket{\varphi_+} = \ket{0} \\
  U(\varphi, \theta)\ket{\varphi_-} = \ket{1}
\end{align*}
and 
\begin{align*}
  U^{-1}(\phi, \theta) \ket{0} = \ket{\psi_+} \\
  U^{-1}(\phi, \theta)\ket{1} = \ket{\psi_-} \\
  U^{-1}(\phi, \theta)\ket{0} = \ket{\varphi_+} \\
  U^{-1}(\varphi, \theta)\ket{1} = \ket{\varphi_-}
\end{align*}
Now, we can define a rotation from

\[\ket{0} \quad \text{to} \quad \cos\left(\frac{\phi}{2} \right) \ket{0} + \sin\left(\frac{\phi}{2} \right) e^{i \theta} \ket{1}  = \ket{\psi_+}\]
and from
\[\ket{1} \quad \text{to} \quad \sin\left(\frac{\phi}{2} \right) \ket{0} + \cos\left(\frac{\phi}{2} \right) e^{i \theta} \ket{1}  = \ket{\psi_-}\]
as follows:
\begin{equation}
  U(\phi, \theta) = 
  \begin{pmatrix}
  e^{i\theta/2} & 0 \\
  0 & e^{-i\theta/2}
  \end{pmatrix} 
  \begin{pmatrix}
  \cos\left(\frac{\phi}{2} \right) & \sin\left(\frac{\phi}{2} \right) \\
  \sin\left(\frac{\phi}{2} \right) & -\cos\left(\frac{\phi}{2} \right)
  \end{pmatrix}
\end{equation}
and the inverse transformation is given by
\begin{equation}
  U^{-1}(\phi, \theta) = 
   \begin{pmatrix}
  e^{-i\theta/2} & 0 \\
  0 & e^{i\theta/2}
  \end{pmatrix}
  \begin{pmatrix}
  \cos\left(\frac{\phi}{2} \right) & \sin\left(\frac{\phi}{2} \right) \\
  \sin\left(\frac{\phi}{2} \right) & -\cos\left(\frac{\phi}{2} \right)
  \end{pmatrix} 
\end{equation}
Now, we can define a quantum circuit that clones the states $\ket{\psi}$ and $\ket{\varphi}$ as follows:
\begin{align*}
  \Qcircuit @C=1em @R=1em {
    & \lstick{\ket{\psi_{\pm}}} & \gate{U(\phi, \theta)} & \ctrl{1}
    & \gate{U^{-1}(\phi, \theta)} & \qw & \rstick{\ket{\psi_{\pm}}} \\
    & \lstick{\ket{0}} & \qw & \targ
    & \gate{U^{-1}(\phi, \theta)} & \qw & \rstick{\ket{\psi_{\pm}}}
  }
\end{align*}
The same follows for the state $\ket{\varphi}$.\\

So far we have shown that a device that can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ can be used to build a device that clones the states $\ket{\psi}$ and $\ket{\varphi}$. But for this we need to know if our qubit is in either 
$\ket{\psi}$ or $\ket{\varphi}$, so we can plug the correct phase and angle into the rotation operator, which is 
in contradiction with the non-cloning theorem.\\

Conversely, given the aforementioned cloning circuit, we can use it to distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ by

\begin{enumerate}
  \item Making n copies of our unknown state $\ket{\chi}$ which is either $\ket{\psi}$ or $\ket{\varphi}$. 
  \item Measuring the $\ket{\chi}$ state in the computational basis $\ket{\psi_{\pm}}$, i.e $\bra{\psi}\ket{\chi}$.
\end{enumerate}
if $\ket{\chi} = \ket{\psi}$, then the outcome is 1, and if $\ket{\chi} = \ket{\varphi}$, then the outcome is $\bra{\psi}\ket{\varphi}$.

\[
\text{Probability} \quad | \bra{\psi}\ket{\psi} |^2 = 1 \quad \text{and probability} \quad | \bra{\psi}\ket{\varphi} |^2 < 1
\]
Let $|\bra{\chi}\ket{\varphi}|^2 = p$, then the probability of measuring an overlap for n measurements of $\bra{\psi}\ket{\phi}$ is  
\[
p^n  = |\bra{\chi}\ket{\varphi}|^{2n}
\]
for example, $\displaystyle{\left \{ \ket{0}, \frac{\ket{0} + \ket{1}}{\sqrt{2}} \right\}}$
\[
p = \left| \bra{0} \left( \frac{\ket{0} + \ket{1}}{\sqrt{2}} \right) \right|^2 = \frac{1}{2}
\]
\[
p^n = \left( \frac{1}{2} \right)^n, \quad \text{and for four measurements} \quad p^4 = \left( \frac{1}{2} \right)^4 = \frac{1}{16}
\]
So we can see that as long as we can make n copies of the state $\ket{\chi}$, we can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$.\\

Even if $\ket{\psi}\sim \ket{\varphi}$, we can still distinguish between them by making enough n copies of the state $\ket{\chi}$ and measuring them in the appropriate 
basis.\\

\end{document}

### 2024-04-23 project/paper-v2.tex
assignments-project

% \documentclass[fleqn,10pt]{wlscirep}
\documentclass[prx,twocolumn,floatfix,superscriptaddress,longbibliography]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{bm}
\usepackage{tikz}
\usepackage{subfiles}
\usepackage{mathtools}
\usetikzlibrary{shapes,arrows,positioning}
\usepackage{standalone}
\usepackage{amssymb,amsmath,amstext}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{appendix}
\usepackage[T1]{fontenc}
\usepackage{bbold}
\usepackage{bbm}
\usepackage{float}
\usepackage{latexsym}
\usepackage{xr-hyper}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=magenta]{hyperref}
\usetikzlibrary{quantikz2}
\usepackage{braket}
\usepackage{physics}
\usepackage{algpseudocode}
\usepackage{placeins}
\usepackage{multirow}
\usepackage{booktabs}
\graphicspath{{./Figures/}}
%\usepackage[latin1]{inputenc}
%\usepackage{hyphenat}

\usepackage{hhline}
\bibliographystyle{apsrev4-2}
\def\bibsection{\section{\refname}} 
\usetikzlibrary{shapes,arrows}
\renewcommand{\vec}[1]{\boldsymbol{#1}}  % Bold vectors instead of arrow vectors
%\newcommand{\<}{\langle}
%\renewcommand{\>}{\rangle}

\long\def\ca#1\cb{} %Use for commenting out: \ca...\cb

\def\cites#1{XXX Need citation #1 XX}

\begin{document}

\title{Quantum Algorithms for Portflolio Optimisation in Finance}

\author{G. Balarezo}
\thanks{e-mail: juan.balarezo@yachaytech.edu.ec}
\affiliation{School of Physical Sciences and Nanotechnology,\\ Yachay Tech University,\\Urcuqui, Ecuador}

\begin{abstract}
Quantum computing has the promise to address computational problems that are considered intractable with our classical computational models. Nevertheless, due to the current limitations of the available quantum hardware, we are still years away from achieving a universal fault-tolerant quantum computer. In the interim, the development of quantum algorithms for specific problems and their implementation on near-term quantum devices is a promising approach. In this work, we review the most promising quantum algorithms for financial applications, focusing on Portfolio Optimisation and Option Pricing. We describe the ideas behind important quantum algorithms such as Variational Quantum Algorithms (VQAs) and Quantum Amplitude Estimation (QAE) and their theoretical performance. We provide an overview of the hardware available for their implementation and discuss the results 
obtained by some study cases, comparing them with their classical counterparts. Finally, we discuss the future prospects of these algorithms in the financial industry.
  \\
  \textbf{Keywords:} Quantum algorithms, Quantum approximate optimization algorithm, Quantum amplitude estimation, Portfolio optimisation, Quantum computing.
\end{abstract}

\maketitle

\section{Introduction}

Quantum computing has become a rapidly growing field of research in recent years, sparking interest from both academic and industrial sectors, including the financial industry in particular \cite{Hassija2020}. The potential of quantum computers to address computational problems that are considered intractable with our classical models of computation \cite{nielsen2010quantum} has driven a considerable amount of research and innovation in the development of Noisy Intermediate-Scale Quantum computers (NISQ) and quantum algorithms capable of leveraging this computational power \cite{Huang2023}.

These near-term devices, with limited qubit capabilities and high error rates, have shown promising results. Implemented along with classical computers and adequate quantum algorithms, they can offer significant speedups for some important but classically inefficiently solvable problems in areas such as Optimisation \cite{Symons2023}, Stochastic Modelling \cite{Herbert2022}, and Machine Learning \cite{Zeguendry2023}. These advancements are particularly relevant to the financial sector \cite{Egger2020a}, where they could yield substantial speedups in tasks such as portfolio optimisation and option pricing \cite{Hassija2020}. These tasks have a high 
computational cost that coupled with growing complexity of the financial markets, the need for real-time analysis of large amounts of data and fast decision-making \cite{book:2139036}, makes quantum computing 
an attractive option for financial institutions.

In the Modern Portfolio Theory proposed by Markowistz \cite{Wilmott2007}, the goal of portfolio optimisation is to find the best combination of assets that maximises the return for a given level of risk and some constraints. This is a highly constrained quadratic optimisation problem which can be formulated in several ways depending on the conditions of the 
investor \cite{Herman2022}. Machine Learning based models (ML) such as the one proposed by Mazumdar et al. \cite{Mazumdar2020} are widely used for minimising risk contributors and maximising asset diversification.

In particular, quantum approaches to portfolio optimisation are gaining significant attention \cite{Quarterly2020}. In 2019 Kerenidis, Prakash and Szilágyi proposed a quantum algorithm for the constrained portfolio optimisation problem. Based on an interior point method, this algorithm reduces the optimisation problem to a second order cone program (SOCP). The authors suggest that their algorithm could achieve a $\mathcal{O} (n)$ speedup over  its classical counterpart.

Other approaches such as the variational quantum algorithms (VQAs) \cite{Yuan2019, Cerezo2021, McClean2016, Amaro2022} work by reformulating the portfolio optimisation problem as a 
quadratic unconstrained binary optimisation (QUBO) problem. Glover et al. \cite{Glover2019} covers some methods to reformulate any optimisation problem as a QUBO problem, and Date et al. \cite{Date2019} proposed an algorithm that allows the efficient embedding of a QUBO problem in a quantum annealer.  Using these concepts, Variational Quantum 
Eigensolver (VQE) \cite{Huang2023, McClean2016, Peruzzo2014, Barkoutsos2020} and Quantum Approximate Optimisation Algorithm (QAOA) \cite{Farhi2014, Blekos2024} have been proposed for portfolio optimisation \cite{Egger2020}. Quantum annealers such as those developed by D-Wave Systems \cite{King2023} are been used for practical implementations of the aforementioned algorithms
\cite{Phillipson2020, Venturelli2019}. Nevertheless, it remains unclear whether these quantum algorithms provide any effective speedup over classical methods for these tasks.

Option pricing is another important task in finance where quantum algorithms have been applied. The implementation of the
Quantum Amplitude Estimation (QAE) algorithm \cite{Woerner2019, Nakaji2020, Stamatopoulos2020} has been proposed for the calculation of the expected value of the option price. This algorithm is
expected to provide a quadratic speedup over the classical Monte Carlo methods used for this task.

The aim of this work is to provide a review of the quantum algorithms for specific financial applications, focusing on Portfolio Optimisation and Option Pricing. Table \ref{Tab:paper_overview} provides an overview of the topics covered in this work, as well as the relevant literature reviewed.

In Sec. \ref{sec:background} we introduce necessary concepts in finance and quantum computing. Then, in Sec. \ref{sec:literature1} and Sec. \ref{sec:literature2}  we present a review of relevant literature on quantum algorithms proposed for the aforementioned financial applications. We also 
  introduce the classical methods used for these tasks and compare them with their quantum counterparts.

In Sec. \ref{sec:hardware} we provide an overview of the hardware implementations of these algorithms. Finally, Sec. \ref{sec:discussion} concludes this work and discusses the future prospects of quantum algorithms in the financial industry.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table*}[t!] 
  \begin{center}
    \caption{\label{Tab:paper_overview} Criteria for the selection of the papers reviewed in this work.}
    \begin{tabular}{c|c|c|c} \hline \hline
\textbf{Metodology} &
  \textbf{Application} &
  \textbf{Algorithm} &
  \textbf{Hardware} \\ \hline
Optimisation &
  \begin{tabular}[c]{@{}c@{}}Portfolio \\ optimisation\end{tabular} &
    \begin{tabular}[c]{@{}c@{}}Approximate \\ Optimisation\\ (SOCP, VQE, QAOA) \\
    \cite{Mazumdar2020}, \cite{Venturelli2019}, \cite{Phillipson2020}, \cite{Rebentrost2018a}  \end{tabular} &
  \begin{tabular}[c]{@{}c@{}}Gate-based quantum computer \\ Quantum Annealer\end{tabular} \\ \hline  
    %\multicolumn{1}{}{  Gate-based quantum computer \\ Quantum Annealer} \\ \hline
Monte Carlo &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Option pricing \end{tabular}} &
  \begin{tabular}[c]{@{}c@{}}Search and Count\\ (QAE, QPA, QPE) \\ 
  \cite{Nakaji2020},\cite{Woerner2019}, \cite{Stamatopoulos2020}, \cite{Rebentrost2022} \cite{Rebentrost2018} \end{tabular} &
\begin{tabular}[c]{@{}c@{}} Gate-based quantum \\ computer\end{tabular} \\ \hline \hline
\end{tabular}
\end{center}
\end{table*}

\section{background}\label{sec:background}

\subsection{Some basics concepts}

\subsubsection{Assets and Portfolio Optimisation}

\subsubsection{Option Pricing}

\subsubsection{Quantum Computing}

\section{Portfolio Optimisation}\label{sec:literature1}

\subsection{Classical Portfolio Optimisation}

\subsection{Quantum Portfolio Optimisation}

\section{Option Pricing}\label{sec:literature2}

\subsection{Classical Option Pricing}

\subsection{Quantum Option Pricing}

\section{Hardware Implementations}\label{sec:hardware}

\section{Discussion and Conclusion}\label{sec:discussion}

\FloatBarrier

\bibliography{./references-2.bib}

\end{document}

### 2024-04-25 assignments/assignment-3.tex
adding assignment 3

--- 2024-04-23 assignments/assignment-2.tex
@@
 \usepackage{subcaption}
-\usepackage{../../../tex-packages/physics}
-\usepackage{../../../tex-packages/qcircuit}
+\usepackage{physics}
+\usepackage{tikz}
+\usetikzlibrary{quantikz2}
+\usepackage{braket}
 \usepackage[framemethod=tikz]{mdframed}
@@
 \usepackage{empheq}
-\usepackage{tikz}
 \usepackage{color}
@@
 \title{Quantum Computation and Quantum Information\\
-Assignment 2}
+Assignment 3}
 \author{Gabriel Balarezo}
@@
 
-Given a device that can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ (without measurement), we can then design a quantum circuit
-that clones the states $\ket{\psi}$ and $\ket{\varphi}$ as we like.\\ 
-First, let's define the non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ as follows:
-\begin{align}
-  \ket{\psi} & \in \{\ket{\psi_+}, \ket{\psi_-}\} \\
-  \ket{\varphi} & \in \{\ket{\varphi_+}, \ket{\varphi_-}\}
-\end{align}
-then, there exists a single-qubit operator, i.e. a rotation such that:
-
-\begin{align*}
-  U(\phi, \theta) \ket{\psi_+} = \ket{0} \\
-  U(\phi, \theta)\ket{\psi_-} = \ket{1} \\
-  U(\phi, \theta)\ket{\varphi_+} = \ket{0} \\
-  U(\varphi, \theta)\ket{\varphi_-} = \ket{1}
-\end{align*}
-and 
-\begin{align*}
-  U^{-1}(\phi, \theta) \ket{0} = \ket{\psi_+} \\
-  U^{-1}(\phi, \theta)\ket{1} = \ket{\psi_-} \\
-  U^{-1}(\phi, \theta)\ket{0} = \ket{\varphi_+} \\
-  U^{-1}(\varphi, \theta)\ket{1} = \ket{\varphi_-}
-\end{align*}
-Now, we can define a rotation from
-
-\[\ket{0} \quad \text{to} \quad \cos\left(\frac{\phi}{2} \right) \ket{0} + \sin\left(\frac{\phi}{2} \right) e^{i \theta} \ket{1}  = \ket{\psi_+}\]
-and from
-\[\ket{1} \quad \text{to} \quad \sin\left(\frac{\phi}{2} \right) \ket{0} + \cos\left(\frac{\phi}{2} \right) e^{i \theta} \ket{1}  = \ket{\psi_-}\]
-as follows:
-\begin{equation}
-  U(\phi, \theta) = 
-  \begin{pmatrix}
-  e^{i\theta/2} & 0 \\
-  0 & e^{-i\theta/2}
-  \end{pmatrix} 
-  \begin{pmatrix}
-  \cos\left(\frac{\phi}{2} \right) & \sin\left(\frac{\phi}{2} \right) \\
-  \sin\left(\frac{\phi}{2} \right) & -\cos\left(\frac{\phi}{2} \right)
-  \end{pmatrix}
-\end{equation}
-and the inverse transformation is given by
-\begin{equation}
-  U^{-1}(\phi, \theta) = 
-   \begin{pmatrix}
-  e^{-i\theta/2} & 0 \\
-  0 & e^{i\theta/2}
-  \end{pmatrix}
-  \begin{pmatrix}
-  \cos\left(\frac{\phi}{2} \right) & \sin\left(\frac{\phi}{2} \right) \\
-  \sin\left(\frac{\phi}{2} \right) & -\cos\left(\frac{\phi}{2} \right)
-  \end{pmatrix} 
-\end{equation}
-Now, we can define a quantum circuit that clones the states $\ket{\psi}$ and $\ket{\varphi}$ as follows:
-\begin{align*}
-  \Qcircuit @C=1em @R=1em {
-    & \lstick{\ket{\psi_{\pm}}} & \gate{U(\phi, \theta)} & \ctrl{1}
-    & \gate{U^{-1}(\phi, \theta)} & \qw & \rstick{\ket{\psi_{\pm}}} \\
-    & \lstick{\ket{0}} & \qw & \targ
-    & \gate{U^{-1}(\phi, \theta)} & \qw & \rstick{\ket{\psi_{\pm}}}
-  }
-\end{align*}
-The same follows for the state $\ket{\varphi}$.\\
-
-So far we have shown that a device that can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ can be used to build a device that clones the states $\ket{\psi}$ and $\ket{\varphi}$.
-But for this we need to know if our qubit is in either
-$\ket{\psi}$ or $\ket{\varphi}$, so we can plug the correct phase and angle into the rotation operator, which is 
-in contradiction with the non-cloning theorem.\\
-
-Conversely, given the aforementioned cloning circuit, we can use it to distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$ by
-
-\begin{enumerate}
-  \item Making n copies of our unknown state $\ket{\chi}$ which is either $\ket{\psi}$ or $\ket{\varphi}$. 
-  \item Measuring the $\ket{\chi}$ state in the computational basis $\ket{\psi_{\pm}}$, i.e $\bra{\psi}\ket{\chi}$.
-\end{enumerate}
-if $\ket{\chi} = \ket{\psi}$, then the outcome is 1, and if $\ket{\chi} = \ket{\varphi}$, then the outcome is $\bra{\psi}\ket{\varphi}$.
-
-\[
-\text{Probability} \quad | \bra{\psi}\ket{\psi} |^2 = 1 \quad \text{and probability} \quad | \bra{\psi}\ket{\varphi} |^2 < 1
-\]
-Let $|\bra{\chi}\ket{\varphi}|^2 = p$, then the probability of measuring an overlap for n measurements of $\bra{\psi}\ket{\phi}$ is  
-\[
-p^n  = |\bra{\chi}\ket{\varphi}|^{2n}
-\]
-for example, $\displaystyle{\left \{ \ket{0}, \frac{\ket{0} + \ket{1}}{\sqrt{2}} \right\}}$
-\[
-p = \left| \bra{0} \left( \frac{\ket{0} + \ket{1}}{\sqrt{2}} \right) \right|^2 = \frac{1}{2}
-\]
-\[
-p^n = \left( \frac{1}{2} \right)^n, \quad \text{and for four measurements} \quad p^4 = \left( \frac{1}{2} \right)^4 = \frac{1}{16}
-\]
-So we can see that as long as we can make n copies of the state $\ket{\chi}$, we can distinguish between non-orthogonal states $\ket{\psi}$ and $\ket{\varphi}$.\\
-
-Even if $\ket{\psi}\sim \ket{\varphi}$, we can still distinguish between them by making enough n copies of the state $\ket{\chi}$ and measuring them in the appropriate 
-basis.\\
-
 \end{document}

### 2024-04-29 lecture-notes/lecture-notes.tex
update .gitignore file

\documentclass[10pt]{article}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage[british]{babel}
\usepackage[utf8]{inputenc}
\usepackage[OT1]{fontenc}
\usepackage{amsfonts, amsmath, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{physics}
\usepackage{empheq}
\usepackage[framemethod=tikz]{mdframed}
\usepackage{bbold}
\usepackage{tikz}
\usetikzlibrary{quantikz2}
\usepackage{braket}
\usepackage{titlesec}
\usepackage{adjustbox}
\usepackage{relsize}
\usepackage{bm}
\usepackage[most]{tcolorbox}
\numberwithin{equation}{section}

\usepackage{cancel}
\usepackage{wrapfig}
\usepackage{float}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\graphicspath{{../fig-files/notes-figs/}}

\hypersetup{
    colorlinks=true,
    citecolor=blue,
    filecolor=blue,
    linkcolor=blue,
    urlcolor=black
}
\newcounter{countCode}
\lstnewenvironment{code} [1][caption=Ponme caption, label=default]{%
	\renewcommand*{\lstlistingname}{Listado} 
	\setcounter{lstlisting}{\value{countCode}} 
	\lstset{ %
	language=java,
	basicstyle=\ttfamily\footnotesize,       % the size of the fonts that are used for the code
	numbers=left,                   % where to put the line-numbers
	numberstyle=\sc,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. 
	numbersep=5pt,                 % how far the line-numbers are from the code
	numberstyle=\color{red!50!blue},
    	backgroundcolor=\color{lightgray!20},
	rulecolor=\color{blue},
	keywordstyle=\color{red}\bfseries,
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	frame=single,                   % adds a frame around the code
	framexleftmargin=0mm,
	numberblanklines=false,
	xleftmargin=5pt,
	breaklines=true,
	breakatwhitespace=true,
	breakautoindent=true,
	captionpos=t,
	texcl=true,
	tabsize=2,                      % sets default tabsize to 3 spaces
	extendedchars=true,
	inputencoding=utf8, 
	escapechar=\%,
	morekeywords={print, println, size, background, strokeWeight, fill, line, rect, ellipse, triangle, arc, save, PI, HALF_PI, QUARTER_PI, TAU, TWO_PI, width, height,},
	emph=[1]{print,println,}, emphstyle=[1]{\color{blue}}, % Mis palabras clave.
	emph=[2]{width,height,}, emphstyle=[2]{\bf\color{violet}}, % Mis palabras clave.
	emph=[3]{PI, HALF_PI, QUARTER_PI, TAU, TWO_PI}, emphstyle=[3]\color{orange!50!violet}, % Mis palabras clave.
	emph=[4]{line, rect, ellipse, triangle, arc,}, emphstyle=[4]\color{green!70!black}, % Mis palabras clave.
	%emph=[5]{size, background, strokeWeight, fill,}, emphstyle=[5]{\tt \color{red!30!blue}}, % Mis palabras clave.
	%emph={[2]sqrt,baset}, emphstyle={[2]\color{blue}}, % f(sqrt(2)), sqrt a nivel 2 se pondrá azul
	#1}}{\addtocounter{countCode}{1}}

%\newtheorem{definition}{Definition}[section]

%\newcounter{fancydefinition}[section]
%\renewcommand{\thefancydefinition}{\thesection.\arabic{fancydefinition}}

%\newenvironment{fancydefinition}[1][]%
%  {\refstepcounter{fancydefinition}%
%\begin{definition}[Definition \thefancydefinition\ #1]}%
%  {\end{definition}}

\newtheoremstyle{defi}
  {\topsep}%
  {\topsep}%
  {\normalfont}%
  {}%
  {\bfseries}% 
  {:}%
  {.5em}%
  {\thmname{#1}\thmnote{~(#3)}}%
\theoremstyle{defi}
\newmdtheoremenv{definitioni}{Definition}[section]
\newmdtheoremenv[
hidealllines=true,
leftline=true,
innertopmargin=0pt,
innerbottommargin=0pt,
linewidth=4pt,
linecolor=gray!40,
innerrightmargin=0pt,
]{definitionii}{Definition}[section]

\newmdtheoremenv[
roundcorner=5pt,
innertopmargin=0pt,
innerbottommargin=5pt,
linewidth=4pt,
linecolor=gray!40,
]{definitioniii}{Definition}[section]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% command for questions 
\newtcolorbox{question}{
freelance,
breakable,
before=\par\vspace{2\bigskipamount}\noindent,
after=\par\bigskip,
frame code={
  \node[
  anchor=south west,
  inner xsep=8pt,
  xshift=8pt,
  rounded corners=5pt,
  font=\bfseries\color{white},
  fill=gray] at (frame.north west) (tit) {\strut Question:};
  \draw[
  line width=3pt,
  rounded corners=5pt,gray
  ] (tit.west) -| (frame.south west) -- ([xshift=15pt]frame.south west);
},
interior code={},
top=2pt
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{apsrev4-1}
%\bibliographystyle{~/Desktop/revtex4-2.bst}
%\bibliographystyle{~/Documents/tex-packages/aps.bst}

\title{\textsf{\textbf{Lecture Notes on Quantum Computing and Quantum Information Theory}}}
\author{J. Gabriel Balarezo}
\date{\today}
\begin{document}
\maketitle \tableofcontents

\begin{abstract}
  In this work we present a set of lecture notes on quantum computing and quantum information theory. The notes are based on the book by Nielsen and Chuang \cite{nielsen2010quantum}. The notes are intended to be used as a guide for a course on quantum computing and quantum information theory. Starting from the fundamentals of quantum computing (quantum mechanics review, definition of a qubit, algorithms and entangelment), this course covers a wide number of interesnting topics in quantum computing and quantum information.

\end{abstract}

\section{Fundamentals of Quantum Computing and Quantum Information}

\emph{Quantum Computing and Quantum Information} is undoubtedly a fascinating area, and is probably the path to follow in order to revolutionise the computational sciences. But before that, we must understand the classical representation of information.

Classically, we represent the information using discrete values, i.e bits. A bit is the smallest unit of information, and can have either a value of $0$ or $1$.  We may say that the classic representation of the information is "deterministic", since we can tell the state (it must be 0 or 1). Similarly, in quantum computing we also have the unit of information. We call it a \emph{qubit}, and unlike the classical bits, a qubit not only can take the discrete values of $0$ and $1$, but a linear combination\footnote{This purely quantum mechanical phenomena is called "superposition".} of them both
\begin{equation}
  \label{eq:1}
  \big|\psi\big>  = \alpha \big|0\big> + \beta \big|1\big>,
\end{equation}

More technically, we can say that a qubit $\big|\psi\big>$ is a superposition of two eigenstates\footnote{Two states {i, j} are said to be eigenstates if $\big<i\big|j\big>=0$.} $\big|0\big>$ and $\big|1\big>$ where $\alpha,\,\, \beta \in \mathbb{C}$ and $\big|\alpha\big|^2 + \big|\beta\big|^2 = 1$. We call this a probabilistic representation of the information.

\subsection{The Bloch Sphere}

The Bloch Sphere is a geometrical representation of the qubit. It is a unit sphere in $\mathbb{R}^3$ where the north and south poles represent the states $\big|0\big>$ and $\big|1\big>$ respectively. The equator of the sphere represents the superposition of the states $\big|0\big>$ and $\big|1\big>$. The states $\big|\psi\big>$ are represented by the points on the surface of the sphere. The Bloch Sphere is a useful tool to visualize the qubit states and their transformations.

\begin{equation}
  \label{eq:2}
  \big|\psi\big> = \cos\left(\frac{\varphi}{2}\right)\big|0\big> + e^{i\theta}\sin\left(\frac{\varphi}{2}\right)\big|1\big>,
\end{equation}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{fig-1-bloch-sphere.pdf}
  \caption{Bloch Sphere representation of a qubit state $\big|\psi\big>$.}
  \label{fig:bloch-sphere}
\end{figure}
where $\theta$ and $\varphi$ are the polar and azimuthal angles of the qubit state $\big|\psi\big>$.

\subsection{Classical Logic Gates}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{logic-gates.pdf}
  \caption{Main logic gates.}
  \label{fig:logic-gates}
\end{figure}

In classical computing, we have the concept of logic gates. These are the basic building blocks of the digital circuits. The most common logic gates are the AND, OR, NOT, NAND, NOR, and XOR gates. In quantum computing, we also have the concept of quantum gates. These are the basic building blocks of the quantum circuits. The most common quantum gates are the Pauli gates, the Hadamard gate, the phase gate, the CNOT gate, and the Toffoli gate.

We say that the NAND gate is universal for classical computing, since any classical circuit can be built using only NAND gates. Similarly, the CNOT gate is universal for quantum computing, since any quantum circuit can be built using only CNOT gates. Even more specific, the NAND Gate is everyting necessary for a "Turin complete" model of computation.

\begin{definitionii}[Turin Complete]
  A model of computation is said to be Turin complete if it can be used to simulate any Turing machine. A Turing machine is a mathematical model of computation that defines an abstract machine that manipulates symbols on a strip of tape according to a table of rules. A Turing machine can simulate any computer algorithm, no matter how complicated.
\end{definitionii}

\subsection{Modelling Quantum Systems}
In order to depict the power of quantum computing, we must first provide an example where the classical computing fails. Let's consider the Many-body problem for atoms.

Consider a system of $N_I$ ions and $N_e$ electrons with ionic charges $eZ_I$, and where the ions are @ $\vec{R}_I$ and the electrons are @ $\vec{r}_i$. The electrons are the particles that we want to simulate. The electrons are fermions, and they obey the Pauli Exclusion Principle. This principle states that no two electrons can occupy the same quantum state. This means that the state of the system is a superposition of all the possible states of the electrons. This is a very hard problem to solve classically, since the number of states grows exponentially with the number of electrons. This is a problem that can be solved using quantum computing.

\begin{definitionii}[Born-Oppenheimer approximation]
The Born-Oppenheimer approximation is a foundational concept in molecular quantum mechanics. It allows for the separation of electronic and nuclear motions in molecules based on the vast difference in mass between electrons and nuclei. This approximation simplifies the Schrödinger equation by treating the electronic structure independently of nuclear motion. It enables the electronic wave function and energy to be solved first, considering the nuclei as fixed. Then, the nuclear motion can be treated semiclassically using the electronic energy as an effective potential. This approximation is crucial for understanding molecular structure and dynamics in computational chemistry and spectroscopy.
\end{definitionii}

We consider the wavefunction to be separable, i.e $\mathlarger{\Psi}(R_1, \dots, R_N; r_1, \dots, r_N)= \mathlarger{\chi}(R_1, \dots, R_N)\mathlarger{\psi}(r_1, \dots, r_N)$.  Where $\mathlarger{\chi}(R_1, \dots, R_N)$ is the nuclear wavefunction and $\mathlarger{\psi}(r_1, \dots, r_N)$ is the electronic wavefunction. Then, the Many-body Hamiltonian has the form

\begin{align}
  \label{eq:3}
  \hat{H} &= \frac{-\hbar^2}{2m_e}\sum_{i=1}^{N_e} \nabla_{\vec{r}_i}^2 + \frac{e^2}{4\pi\epsilon_0}\sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\frac{1}{|\vec{r}_i - \vec{r}_j|} - 
  \frac{e^2}{4\pi\epsilon_0}\sum_{I=1}^{N_I}Z_I\sum_{i=1}^{N_e}\frac{1}{|\vec{r}_i - \vec{R}_I|} + \nonumber\\ 
  &- \frac{-\hbar^2}{2M_I}\sum_{I=1}^{N_I} \nabla_{\vec{R}_I}^2 + \frac{e^2}{4\pi\epsilon_0}\sum_{I=1}^{N_I}\sum_{J=I+1}^{N_I}\frac{Z_IZ_J}{|\vec{R}_I - \vec{R}_J|} 
\end{align}

If $\mathlarger{\Psi}(\vec{R}_1,\dots, \vec{R}_{N_I}; \vec{r}_1,\dots, \vec{r}_{N_e})$ is the full many-body wavefunction, then the Schrödinger equation for the system is

\begin{equation}
  \label{eq:4}
  \hat{H}\mathlarger{\Psi} = \mathlarger{\chi}(R_1, \dots, R_N) \hat{H}_e\mathlarger{\psi}^{\vec{R}_1, \dots, \vec{R}_{N_I}}(r_1, \dots, r_{N_e})
  + \mathlarger{\psi}^{\vec{R}_1, \dots, \vec{R}_{N_I}}(r_1, \dots, r_{N_e})\hat{H}_n\mathlarger{\chi}(R_1, \dots, R_{N_I})
  \end{equation}
 where $\chi$ and $\psi$  are the ionic and electronic components of $\mathlarger{\Psi}$.

\begin{mdframed}[backgroundcolor = gray!30,
  frametitle = Idea, roundcorner = 10pt]
  The electrons will relax to their ground state before the ionic core can be disturbed, so that wecan treat the system as classical ions with quantum mechanical electrons.
\end{mdframed}

\begin{equation*}
  \nabla^2 = \frac{\partial^2 f}{\partial x^2} + \frac{\partial^2 f}{\partial y^2} + \frac{\partial^2 f}{\partial z^2} \rightarrow \text{choose}\quad 
  \frac{\partial^2 f}{\partial x^2}  = \frac{\partial}{\partial x}\left(\frac{\partial f}{\partial x}\right)
\end{equation*}
writing the numerical approximation
\begin{equation}
  \label{eq:5}
  \frac{\partial^2 f}{\partial x^2} \approx \frac{f(x+h) - 2f(x) + f(x-h)}{h^2}
\end{equation}
and its matrix representation
\begin{equation}
  \label{eq:6}
  \nabla^2 f \approx \frac{1}{h^2}\begin{pmatrix}
    -2 & 1 & 0 & \dots & 0\\
    1 & -2 & 1 & \dots & 0\\
    0 & 1 & -2 & \dots & 0\\
    \vdots & \vdots & \vdots & \ddots & \vdots\\
    0 & 0 & 0 & \dots & -2
  \end{pmatrix} 
  \begin{pmatrix}
    f(x_0)\\
    f(x_1)\\
    f(x_2)\\
    \vdots\\
    f(x_n)
  \end{pmatrix}
\end{equation}
The electron-electron interaction is a two-point function. Now, let's compute the total energy of the system

\begin{align*}
  E_{\text{total}} &= E\big<\Psi\big|\Psi\big> = \big<\Psi\big|\hat{H}\big|\Psi\big>\\
&= \big<\chi\big|\chi\big>\big<\psi\big|\hat{H}_e\big|\psi\big> + \big<\psi\big|\psi\big>\big<\chi\big|\hat{H}_I\big|\chi\big>\\
  &= \big<\psi\big|\hat{T}_e\big|\psi\big> + \big<\psi\big|\hat{V}_{ee}\big|\psi\big> + \big<\psi\big|\hat{V}_{eI}\big|\psi\big> + \big<\chi\big|\hat{T}_I\big|\chi\big> + \big<\chi\big|\hat{V}_{II}\big|\chi\big> 
\end{align*}
therefore

\begin{align}
  \label{eq:7}
  E(\vec{R}_1, \dots, \vec{R}_{N_I}) &= \big<\psi\big|\hat{T}_e\big|\psi\big> + \big<\psi\big|\hat{V}_{ee}\big|\psi\big> + \big<\psi\big|\hat{V}_{eI}\big|\psi\big>\nonumber\\
  &+ T_I(\vec{R}_1, \dots, \vec{R}_{N_I}) + V_{II}(\vec{R}_1, \dots, \vec{R}_{N_I})
\end{align}

Now, write the electron-electron contribution to the total energy

\begin{align}
  \label{eq:8}
  \big<\psi(\vec{r}_1, \dots, \vec{r}_{N_e})\big|\sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\frac{1}{|\vec{r}_i - \vec{r}_j|}\big|\psi(\vec{r}_1, \dots, \vec{r}_{N_e})\big> &= \iiint d^3 \vec{r}_{1}\dots\iiint d^3\vec{r}_{N_e}
  \sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\frac{\big|\psi(\vec{r}_1, \dots, \vec{r}_{N_e})\big|^2}{|\vec{r}_i - \vec{r}_j|}\nonumber\\
  &= \sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\int d^3\vec{r}_{1}\int d^3\vec{r}_{N_e}\frac{\big|\psi(\vec{r}_1, \dots, \vec{r}_{N_e})\big|^2}{|\vec{r}_i - \vec{r}_j|}\nonumber\\
  &= \sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\int d^3\vec{r}_{1} \frac{\rho(\vec{r}_i, \vec{r}_{j})}{\big|\vec{r}_i - \vec{r}_j\big|}\nonumber\\
  &= \sum_{i=1}^{N_e}\sum_{j=i+1}^{N_e}\iint \frac{\rho(\vec{r'}) d^3\vec{r'}}{\big|\vec{r} - \vec{r'}\big|} \rho(\vec{r}) d^3\vec{r} + \int \frac{\delta E_{\text{ex}} \rho(\vec{r}) d^3\vec{r}}{\delta \rho}
\end{align}

The last term in Eq.\eqref{eq:8} is the exchange-correlation energy. The exchange energy is the energy associated with the Pauli exclusion principle. The correlation energy is the energy associated with the electron-electron interaction. 
Computing this term classically is very hard, since the number of states grows exponentially with the number of electrons. 
But, since quantum computing and quantum information is based on long-range correlations, we can compute this term using quantum computing.

\begin{question}
  How many bits of information are needed to describe an oxigen atom's many-body electronic wavefunction $\mathlarger{\psi}(r_1, \dots, r_{N_e})$?
\end{question}
\begin{itemize}
  \item Use a grid representation with a spacing of $h = 0.2$ \AA.
  \item Make each side of the box $L = 10$ \AA $= 1$ nm $= 50$ grid points. 
\end{itemize}

  \begin{equation*}
    \parbox{0.17\textwidth}{number of points} = \left(\frac{L}{h}\right)^3 = 50^3 \quad \text{points}
  \end{equation*}
\begin{equation*}
  \parbox{0.3\textwidth}{number of functional evaluations} = \left(\parbox{0.2\textwidth}{number of grid points}\right)^{N_e} = \left(\frac{L}{h}  \right)^{3\times N_e}
\end{equation*}
grows exponentially with the number of electrons.

\begin{equation*}
 \# \text{data points} = 50^{24} \approx 5 \times 10^{24} = 5\times N_A
\end{equation*}

We can see that we will need the equivalent to $5$ Avogadro's numbers of bits to describe the many-body electronic wavefunction of an Oxigen atom.

\subsection{Quantum Logic Gates}
As aforementioned, for a system of computation to be "Turing complete" it needed only to implement the NAND gate. 
\begin{question}
  Can a quantum system of computation implement a NAND  gate?
\end{question}
Quantum systems obey time reversal symmetry (CAVEAT: this is not true for all quantum systems). This means that the time evolution of a quantum system is reversible. In order to construct the quantum logic gates we use quantum operators\footnote{Any quantum operator $\hat{\theta}$ is time reversible since it has an inverse.}.

So, we need to use "ancillary bits", i.e extra bits with a set initial value to "store" the result
% draw in latex a toffoli gate with states a and b and ancillary bit c = 0

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{toffoli-gate.pdf}
  \caption{Toffoli gate with ancillary bit 0.}
  \label{fig:toffoli-gate}
\end{figure}

The Toffoli gate provides a time reversible, i.e a invertable model for computation which is also "Turing complete".

\begin{definitionii}[Efficiency]
  Time it takes to perform an algorithm $t(n)$ where n is the size of the input (space) is equal to a polynomial function
  \begin{equation*}
    t(n) = P(n) = \mathcal{O}(n^p), \quad p \in \mathbb{N}
  \end{equation*}
  or equivalently
\begin{equation*}
  \exists\, \text{M} \in \mathbb{R} \quad \text{such that} \quad t(n) \leq \text{M}n^p
\end{equation*}
  i.e $t(n)$ is bounded by $n^p$.
\end{definitionii}
\begin{definitionii}[Inefficient]
An algorithm is said to be inefficient if the time it takes grows exponentially as the size of the input, i.e,
\begin{equation*}
  t(n) = 2^n = \mathcal{O}(2^n)
\end{equation*}
\end{definitionii}

\subsection{Requirements for Quantum Computing}
So far, we have learned some definitions and a set of rules for a quantum system of computation to be "Turing complete". Now, we can state the requirements for a quantum system of computation.

\begin{enumerate}
\item Ability to prepare qubits in a well-defined initial input state, i.e  prepare either $\big|0\big>$ or $\big|1\big>$.
\item Ability to read the result of our algorithm, i.e measure the state of the qubit.
  \item Ability to prepare pure quantum superpositions, $\big|\psi\big> = \cos\left(\frac{\varphi}{2}\right)\big|0\big> + e^{i\theta}\sin\left(\frac{\varphi}{2}\right)\big|1\big>$.
  \item Toffoli\footnote{The 3-qubit Toffoli gate may be implemented using a combination of 2-qubit CNOT-gates.} gate implementation -two or three qubit gate.
\end{enumerate}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{cnot-gate.pdf}
  \caption{(a) represents the CNOT gate applied to the qubits a and b. (b) shows the matrix representation of the CNOT gate, and (c) shows its truth table.}
  \label{fig:quantum-computer}
\end{figure}

A single qubit (Eq. \eqref{eq:2}) can be represented as a position on the surface of the unit sphere (see Fig.\ref{fig:bloch-sphere}). Any combination of two qubits takes the form

\begin{equation}
  \label{eq:9}
  \big|\psi\big> = \alpha_{00} \big|00\big> + \alpha_{01} \big|01\big> + \alpha_{10} \big|10\big> + \alpha_{11} \big|11\big>
  \end{equation}

  where $\displaystyle{\sum_{i,j \in\{0, 1\}^2 }|\alpha_{ij}|^2 = 1}$.

\subsection{Quantum Logic Gates}
\begin{mdframed}[backgroundcolor = gray!30,
  frametitle = Goal, roundcorner = 10pt]
  Can we make from the basis states $\big|0\big>$ or $\big|1\big>$ to the superposition of states, i.e onto the circle in the xy axis (frame) of the Bloch sphere?
  \end{mdframed}

  The answer is yes, we can. We can use the Hadamard gate to make the transformation.
\subsubsection{The Hadamard Gate}
The Hadamard gate creates superposition by putting the qubit $\big|0\big>$ and $\big|1\big>$  into a state where there is an equal probability of measuring 0 or 1. It's an essential gate in many quantum algorithms, including quantum teleportation and Grover's search algorithm.
  \begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{hadamard-gate.pdf}
    \caption{Hadamard gate.}
    \label{fig:hadamard-gate}
  \end{figure}
If we apply the Hadamard gate to the qubit $\alpha |0\rangle + \beta |1\rangle$, we get

\begin{equation}
  \label{eq:10}
  \frac{1}{\sqrt{2}} \begin{pmatrix}
    1 & 1\\
    1 & -1
  \end{pmatrix}
  \begin{pmatrix}
    \alpha\\
    \beta
  \end{pmatrix}
  = \frac{\alpha + \beta}{\sqrt{2}}|0\rangle + \frac{\alpha - \beta}{\sqrt{2}}|1\rangle
\end{equation}
Similarly, we can apply the Hadamard gate to the pure superposition of states $\displaystyle{\frac{|0\rangle + i|1\rangle}{\sqrt{2}}}$,

\begin{align}
  \label{eq:11}
  \frac{1}{\sqrt{2}}\frac{1}{\sqrt{2}} \begin{pmatrix}
    1 & 1\\
    1 & -1
  \end{pmatrix}
  \begin{pmatrix}
    1\\
    i
  \end{pmatrix}
  =  \frac{1}{2}\begin{pmatrix}
    1 + i\\
    1 - i
  \end{pmatrix}
\end{align}
consider $(1 + i)$ and $(1 - i)$ as the complex numbers $z_1$ and $z_2$ respectively. Then, we can write the complex numbers as
$z_1 = \sqrt{2}e^{i\frac{\pi}{4}}$ and $z_2 = \sqrt{2}e^{-i\frac{\pi}{4}}$. So we can write the result in Eq.\eqref{eq:11} as

\tikzset{operator/.append style={fill=red!20}}

\begin{equation}
  \label{eq:12}
  \frac{|0\rangle + i|1\rangle}{\sqrt{2}}
  \begin{quantikz}[column sep=0.5cm, thin lines]
  & \gate{H} & 
  \end{quantikz}  
   = \frac{\sqrt{2}\exp(i\pi/4)|0\rangle +\sqrt{2} \exp(-i\pi/4)|1\rangle}{2}
\end{equation}
we can also multiply this state by a phase factor, i.e $\exp(i\pi/4)$, which is a global phase factor. This means it does not change the state of the qubit

\begin{equation*}
  \frac{\sqrt{2}\exp(i\pi/4)|0\rangle +\sqrt{2} \exp(-i\pi/4)|1\rangle}{2}  \exp(i\pi/4) = \frac{|0\rangle - i|1\rangle}{\sqrt{2}} 
\end{equation*}

\begin{equation}
  \label{eq:13}
  \frac{|0\rangle + i|1\rangle}{\sqrt{2}} 
  \begin{quantikz}[column sep=0.5cm, thin lines]
  & \gate{H} &
  \end{quantikz}
   = \frac{|0\rangle - i|1\rangle}{\sqrt{2}}
\end{equation}

\subsubsection{The $\pi/8$ Gate} 
The $\frac{\pi}{8}$ gate introduces a phase shift of $\frac{\pi}{4}$ (or $\frac{\pi}{8}$ radians) to the state $|1\rangle$, leaving $|0\rangle$ unchanged. Mathematically, it applies the transformation:

\begin{equation}
  \label{eq:14}
  \begin{quantikz}[column sep=0.5cm, thin lines]
    & \gate{T} &
  \end{quantikz}
  = \begin{pmatrix}
    1 & 0\\
    0 & e^{i\pi/4}
   \end{pmatrix}
\end{equation}

\subsubsection{Phase Gate}
The phase gate introduces a phase shift of $\frac{\pi}{2}$ (or $\frac{\pi}{2}$ radians) to the state $|1\rangle$, leaving $|0\rangle$ unchanged. Mathematically, it applies the transformation:

\begin{equation}
  \label{eq:15}
  \begin{quantikz}[column sep=0.5cm, thin lines]
    & \gate{S} &
  \end{quantikz}
  = \begin{pmatrix}
    1 & 0\\ 
    0 & i\\ 
    \end{pmatrix}
    \equiv 
    \begin{quantikz}[column sep=0.5cm, thin lines]
      & \gate{T} & \gate{T} &
    \end{quantikz}
\end{equation}
\subsubsection{Pauli-Z Gate}

\begin{equation}
  \label{eq:16}
  \begin{quantikz}[column sep=0.5cm, thin lines]
    & \gate{Z} &
  \end{quantikz}
  = \begin{pmatrix}
    1 & 0\\ 
    0 & -1\\ 
    \end{pmatrix}
\end{equation}
The Pauli-Z gate generates a half-turn in the Bloch sphere about the $z$ axis. With respect to the computational basis, the $Z$ gate flips the phase of the $|1\rangle$ state relative to the $|0\rangle$ state.

\subsubsection{Pauli-X Gate}
\begin{equation}
  \label{eq:17}
  \begin{quantikz}[column sep=0.5cm, thin lines]
    & \gate{X} &
  \end{quantikz}
  = \begin{pmatrix}
    0 & 1\\ 
    1 & 0\\ 
    \end{pmatrix}
\end{equation}
The X-gate is also known as a bit-flip gate. It generates a half-turn in the Bloch sphere about the $x$ axis.

With respect to the computational basis, the $X$ gate is equivalent to a classical NOT operation, or logical negation. The computation basis states are interchanged, so that $|0\rangle$ becomes $|1\rangle$ and $|1\rangle$ becomes $|0\rangle$.

\subsubsection{Pauli-Y Gate}

\begin{equation}
  \label{eq:18}
  \begin{quantikz}[column sep=0.5cm, thin lines]
    & \gate{Y} &
  \end{quantikz}
  = \begin{pmatrix}
    0 & -i\\ 
    i & 0\\ 
    \end{pmatrix}
\end{equation}

The Pauli-Y gate generates a half-turn in the Bloch sphere about the $y$ axis. This gate can be thought as a combination of $X$ and $Z$ gates, $Y = -iZX$. With respect to the computational basis, we interchange the zero and one states and apply a relative phase flip.

The Hadamard Gate, $\pi/8$ Gate and CNOT gate are called universal gates since they are sufficient to implement quantum computation.

\subsection{Bell's States}
The Bell states, named after physicist John S. Bell, are a set of four maximally entangled quantum states of two qubits. They represent fundamental examples of quantum entanglement, wherein the states of two particles become interdependent regardless of the distance between them.

These states, denoted as $\beta_{00}$, $\beta_{01}$, $\beta_{10}$, and $\beta_{11}$, demonstrate distinct patterns of correlation between the qubits, characterized by specific superpositions of the basis states. For instance, the $\beta_{00}$ state represents an equal superposition of both qubits being in the state $|0\rangle$ or $|1\rangle$, while the $\beta_{01}$ state has opposite signs in the superposition.

The importance of Bell states lies in their pivotal role in various quantum information processing tasks, including quantum teleportation, superdense coding, and quantum cryptography. They serve as essential resources for enabling secure communication, efficient information transmission, and the implementation of quantum algorithms. The general diagram for Bell's states  is here

% write the bell 00 state using quantikz 
\begin{equation}
  \label{eq:19}
  \begin{quantikz} 
    \ket{i} \gategroup[wires=2, steps=1, style={dashed, rounded corners,fill = blue!20, inner sep=3pt},background, label style={label position=below, yshift=-0.5cm}]{pure states}& \gate{H} & \ctrl{1} &\rstick[2]{$\ket{B_{ij}}$, $i,j$ $\in$ \{0,1  \} } \\
     \ket{j} &          & \targ{}  & 
  \end{quantikz}
\end{equation}
even more generallly, we can write the Bell states as

\begin{equation}
  \label{eq:20}
  \ket{B_{ij}}  =  
  \begin{quantikz} 
    \lstick[2]{$\ket{B_{00}}$} &  & & \\ 
    & \gate{X^{i}}  & \gate{Z^j} &
  \end{quantikz}
\end{equation}
\newpage
\begin{mdframed}[backgroundcolor = gray!30,
  frametitle = Bell's states, roundcorner = 10pt]
\begin{equation}
  \label{eq:21}
  %draw the bell 00 state using quantikz 
  \begin{quantikz} 
    \lstick{$\ket{0}$} & \gate{H} & \ctrl{1} & \rstick[2]{$\frac{1}{\sqrt{2}}(\ket{00} + \ket{11})$} \\
    \lstick{$\ket{0}$} & \qw & \targ{} &
  \end{quantikz}
  = \ket{B_{00}}
\end{equation}
\begin{equation}
  \label{eq:22}
  %draw the bell 01 state using quantikz2 
  \begin{quantikz} 
    \lstick{$\ket{0}$} & \gate{H} & \ctrl{1} & \rstick[2]{$\frac{1}{\sqrt{2}}(\ket{00} - \ket{11})$} \\
    \lstick{$\ket{1}$} & \qw & \targ{} &
  \end{quantikz}
  = \ket{B_{01}}
\end{equation}
\begin{equation}
  \label{eq:23}
  %draw the bell 10 state using quantikz3
  \begin{quantikz} 
    \lstick{$\ket{1}$} & \gate{H} & \ctrl{1} & \rstick[2]{$\frac{1}{\sqrt{2}}(\ket{10} + \ket{01})$} \\
    \lstick{$\ket{0}$} & \qw & \targ{} &
  \end{quantikz}
  = \ket{B_{10}}
\end{equation}
\begin{equation}
  \label{eq:24}
  %draw the bell 11 state using quantikz4
  \begin{quantikz} 
    \lstick{$\ket{1}$} & \gate{H} & \ctrl{1} & \rstick[2]{$\frac{1}{\sqrt{2}}(\ket{10} - \ket{01})$} \\
    \lstick{$\ket{1}$} & \qw & \targ{} &
  \end{quantikz}
  = \ket{B_{11}}
\end{equation}
  \end{mdframed}

\subsection{Quantum Teleportation}

Quantum teleportation is a process by which the state of a qubit can be transmitted from one location to another, with the help of two classical bits and an entangled pair of qubits. The process involves three parties: the sender, the receiver, and the qubit to be teleported. 
For the matter of this example, let's consider the state of the qubit to be teleported as $\ket{\psi} = \alpha\ket{0} + \beta\ket{1}$. The sender and receiver share an entangled pair of qubits in the Bell state $\ket{B_{00}} = \frac{1}{\sqrt{2}}(\ket{00} + \ket{11})$. The sender performs a Bell measurement on the qubit to be teleported and their half of the entangled pair. The result of the Bell measurement is two classical bits, which are sent to the receiver. The receiver then applies a series of quantum gates based on the classical bits received, to transform their half of the entangled pair into the original state of the qubit to be teleported.

% draw the quantum teleportation circuit using quantikz2 
\begin{equation}
  \label{eq:25}
\begin{quantikz}
  \lstick[2]{Alice} \ket{\psi} && \ctrl{1} & \gate{H} & \meter{M_1} & \setwiretype{c} & \cwbend{2} \\
 \lstick[2]{$\ket{B_{00}}$}    && \targ{}  &         & \meter{M_2}  & \cwbend{1} \setwiretype{c}  \\
\lstick{Bob}           &&          &         &              & \gate{X^{M_2}}   & \gate{Z^{M_1}} 
\end{quantikz}
\end{equation}

Now, let's do the math behind the quantum teleportation. We start with the state of the qubit to be teleported $\ket{\psi} = \alpha\ket{0} + \beta\ket{1}$ and the entangled pair in the Bell state $\ket{B_{00}} = \frac{1}{\sqrt{2}}(\ket{00} + \ket{11})$. The total state of the system is given by

\begin{equation}
  \label{eq:26}
  \ket{\psi} \otimes \ket{B_{00}} = \frac{1}{\sqrt{2}}(\alpha\ket{0} + \beta\ket{1})\otimes(\ket{00} + \ket{11})
\end{equation}

Expanding the tensor product, we get

\begin{equation}
  \label{eq:27}
  \frac{1}{\sqrt{2}}(\alpha\ket{0}\otimes\ket{00} + \alpha\ket{0}\otimes\ket{11} + \beta\ket{1}\otimes\ket{00} + \beta\ket{1}\otimes\ket{11})
\end{equation}

Now, we apply the Hadamard gate to the first qubit of the qubit to be teleported and the CNOT gate to the two qubits. The state of the system after these operations is

\begin{equation}
  \label{eq:28}
  \frac{1}{2}(\alpha\ket{0}\otimes(\ket{0} + \ket{1}) + \alpha\ket{1}\otimes(\ket{0} - \ket{1}) + \beta\ket{0}\otimes(\ket{0} + \ket{1}) + \beta\ket{1}\otimes(\ket{0} - \ket{1}))
\end{equation}

Now, we measure the two qubits and send the results to the receiver. The receiver then applies the Pauli-X and Pauli-Z gates to their qubit based on the classical bits received. The state of the receiver's qubit after these operations is

\begin{equation}
  \label{eq:29}
  \frac{1}{2}(\alpha\ket{0} + \beta\ket{1})
\end{equation}

This is the original state of the qubit to be teleported, $\ket{\psi} = \alpha\ket{0} + \beta\ket{1}$. Thus, the quantum teleportation process has successfully transmitted the state of the qubit from the sender to the receiver.

\newpage 
\bibliography{references-2}
%\bibliography{~/Library/texmf/bibtex/bib/zotero}

\end{document}

### 2024-05-01 assignments/assignment-3.tex
adding assigment 3

@@
 \begin{mdframed}[backgroundcolor = gray!30,
-  frametitle = Exercise 1.2]
-  Explain how a device which, upon input of one of two non-orthogonal quantum states $\ket{\psi}$ and $\ket{\varphi}$ correctly identified the state, could be used to build a device which cloned the states $\ket{\psi}$ and $\ket{\varphi}$, in violation of the non-cloning theorem.
-Conversely, explain how a device for cloning could be used to distinguish non-orthogonal quantum states.
+  frametitle = Exercise 4.30]
+  Suppose $U$ is a single-qubit unitary operation.
+Find a circuit containing $\mathcal{O}(n^2)$ Toffoli, CNOT and single qubit gates which implements a $C^n (U)$ gate (for $n>3$), using no work qubits.
 \end{mdframed}

### 2024-05-03 assignments/assignment-3.tex
finished assignment-3

@@
 % Packages for mathematical symbols and equations
+\usepackage[numbers]{natbib}
 \usepackage{amsmath}
@@
 
+\bibliographystyle{apsrev4-2}
 % Set page margins
@@
 
+So, the circuit we want to implement is the following:
+\begin{equation*}
+  \begin{quantikz}
+    \lstick{$\ket{x_1}$} &  & \ctrl{1} &\\ 
+    \lstick{$\ket{x_2}$} &  & \ctrl{1} & \\
+    \lstick{$\ket{x_3}$} &  & \ctrl{2} & \\
+    \vdots \\
+    \lstick{$\ket{x_n}$} &  & \ctrl{1} & \\
+    \lstick{$\ket{\psi}$} &  & \gate{U} &
+    \end{quantikz}
+\end{equation*}
+
+For this, we will Lemma 6.1 from Ref. \cite{Barenco_1995}, which states that for any unitary $2\times 2$ matrix $U$, a $C^2(U)$ gate can be simulated by the following 
+circuit:
+
+\begin{equation*}
+  \begin{quantikz}
+    & \ctrl{1} & \midstick[3, brackets = none]{=} & & \ctrl{1} & & \ctrl{1} & \ctrl{2} &\\ 
+    & \ctrl{1} & & \ctrl{1} & \targ{} & \ctrl{1} & \targ{} & &  \\
+    & \gate{U} & & \gate{U^{1/2}} &         & \gate{(U^{1/2})^\dagger} &  & \gate{U^{1/2}} &
+    \end{quantikz}
+\end{equation*}
+
+Proof of this lemma can be found in the same reference.\\
+
+A generalisation of Lemma 6.1 can be found in Lemma 7.5 of the same reference, which states that for any unitary $2\times 2$ matrix $U$, a $C^{n-1}(U)$ gate can be simulated by the following circuit (illustrated for $n = 9$)
+
+\begin{equation*} 
+  \begin{quantikz}
+    & \ctrl{1} & \midstick[10, brackets = none]{=} & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\ 
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\ 
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\ 
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\ 
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{1} &\\  
+    & \ctrl{1} &                                  & & \ctrl{1} & & \ctrl{1} & \ctrl{2} &\\ 
+    & \ctrl{1} &                                  &\ctrl{1}&\targ{}&\ctrl{1}&\targ{}& &\\ 
+    & \gate{U} &                                  &\gate{V}&       &\gate{V^\dagger}&       & \gate{V} &
+  \end{quantikz}
+\end{equation*}
+
+given $V$ a unitary matrix such that $V^2 = U$. Another version of this implementation can be found in Ref. \cite{liu2007analytic}
+
+What we want to show is that the circuit above can be implemented using $\mathcal{O}(n^2)$ basic operations.
+
+We can notice that this circuit is a recursive implementation of Lemma 7.5.
+Let $C_{n-1}$ denote the cost of implementing a $C^{n-1}(U)$.
+We can also notice that the cost of applying $C^{1}(V)$ and $C^{1}(V^\dagger)$ gates is $\mathcal{O}(1)$.
+Corollary 7.4 from Ref.\cite{Barenco_1995} states that for an n-bit circuit (for $n\geq 7$), a $C^{n-2}(X)$ gate can be simulated using $\mathcal{O}(n)$ basic operations.
+The cost of implementing $C^{n-2}(V)$ gate is $C_{n-2}$ (by recursion).
+This implies that the cost of implementing the circuit above is $C_{n-1} = C_{n-2} + \mathcal{O}(n)$. This is a linear recurrence relation, and its solution is $C_{n-1} = \mathcal{O}(n^2)$.\\
+
+Note: Illustrations were made using the package \texttt{quantikz2}.
+
+\bibliography{/home/pacman/Documents/git-repositories/quantum-computing/lecture-notes/references-2.bib}
+
 \end{document}

### 2024-05-09 project/paper-v2.tex
update .gitignore file
update paper
update project paper

@@
 In the Modern Portfolio Theory proposed by Markowistz \cite{Wilmott2007}, the goal of portfolio optimisation is to find the best combination of assets that maximises the return for a given level of risk and some constraints.
-This is a highly constrained quadratic optimisation problem which can be formulated in several ways depending on the conditions of the
+This is a highly constrained quadratic optimisation problem, known to be NP-Hard, which can be formulated in several ways depending on the conditions of the
 investor \cite{Herman2022}.
@@
 expected to provide a quadratic speedup over the classical Monte Carlo methods used for this task.
-
-The aim of this work is to provide a review of the quantum algorithms for specific financial applications, focusing on Portfolio Optimisation and Option Pricing.
-Table \ref{Tab:paper_overview} provides an overview of the topics covered in this work, as well as the relevant literature reviewed.
-
-In Sec.
-\ref{sec:background} we introduce necessary concepts in finance and quantum computing.
-Then, in Sec.
-\ref{sec:literature1} and Sec.
-\ref{sec:literature2}  we present a review of relevant literature on quantum algorithms proposed for the aforementioned financial applications.
-We also
-  introduce the classical methods used for these tasks and compare them with their quantum counterparts.
-
-In Sec.
-\ref{sec:hardware} we provide an overview of the hardware implementations of these algorithms.
-Finally, Sec.
-\ref{sec:discussion} concludes this work and discusses the future prospects of quantum algorithms in the financial industry.
 
@@
 \end{table*}
+
+The aim of this work is to provide a review of the quantum algorithms for specific financial applications, focusing on Portfolio Optimisation and Option Pricing.
+Table \ref{Tab:paper_overview} provides an overview of the topics covered in this work, as well as the relevant literature reviewed.
+
+In Sec.
+\ref{sec:background} we introduce necessary concepts in finance and quantum computing.
+Then, in Sec.
+\ref{sec:literature1} and Sec.
+\ref{sec:literature2}  we present a review of relevant literature on quantum algorithms proposed for the aforementioned financial applications.
+We also
+  introduce the classical methods used for these tasks and compare them with their quantum counterparts.
+
+In Sec.
+\ref{sec:hardware} we provide an overview of the hardware implementations of these algorithms.
+Finally, Sec.
+\ref{sec:discussion} concludes this work and discusses the future prospects of quantum algorithms in the financial industry.
+
+Although this work is focused on two financial applications, the quantum algorithms reviewed here can be applied to a wide range of financial problems.
+In recent years, important surveys have been published on the topic of quantum computing in finance \cite{Orus2019,Egger2020a,Albareti2022,Herman2022,Herman2023}
+and we refer the interested readers to these works for a more comprehensive overview of the field.
 
@@
 
-\bibliography{./references-2.bib}
+\bibliography{/home/pacman/Documents/git-repositories/quantum-computing/lecture-notes/bibliography.bib}
 

### 2024-05-26 project/paper-v2.tex
update paper-v2.pdf
paper update

@@
 
-In the Modern Portfolio Theory proposed by Markowistz \cite{Wilmott2007}, the goal of portfolio optimisation is to find the best combination of assets that maximises the return for a given level of risk and some constraints.
+In the Modern Portfolio Theory (MPT) proposed by Markowistz \cite{Markowitz1952}, the goal of portfolio optimisation is to find the best combination of assets that maximises the return for a given level of risk and some constraints.
 This is a highly constrained quadratic optimisation problem, known to be NP-Hard, which can be formulated in several ways depending on the conditions of the
@@
 Using these concepts, Variational Quantum
-Eigensolver (VQE) \cite{Huang2023, McClean2016, Peruzzo2014, Barkoutsos2020} and Quantum Approximate Optimisation Algorithm (QAOA) \cite{Farhi2014, Blekos2024} have been proposed for portfolio optimisation \cite{Egger2020}.
+Eigensolver (VQE) \cite{Huang2023, McClean2016, Peruzzo2014, Barkoutsos2020, Buonaiuto2023} and Quantum Approximate Optimisation Algorithm (QAOA) \cite{Farhi2014, Blekos2024} have been proposed for portfolio optimisation \cite{Egger2020}.
 Quantum annealers such as those developed by D-Wave Systems \cite{King2023} are been used for practical implementations of the aforementioned algorithms
-\cite{Phillipson2020, Venturelli2019}. Nevertheless, it remains unclear whether these quantum algorithms provide any effective speedup over classical methods for these tasks.
+\cite{Yarkoni2022, Phillipson2020, Venturelli2019}. Nevertheless, it remains unclear whether these quantum algorithms provide any effective speedup over classical methods for these tasks.
 
 Option pricing is another important task in finance where quantum algorithms have been applied. The implementation of the
-Quantum Amplitude Estimation (QAE) algorithm \cite{Woerner2019, Nakaji2020, Stamatopoulos2020} has been proposed for the calculation of the expected value of the option price. This algorithm is
-expected to provide a quadratic speedup over the classical Monte Carlo methods used for this task.
+Quantum Amplitude Estimation (QAE) algorithm \cite{Woerner2019, Nakaji2020, Stamatopoulos2020, Rebentrost2018} has been proposed for the calculation of the expected value of the option price.
+This algorithm is
 
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
 \begin{table*}[t!] 
@@
     \begin{tabular}[c]{@{}c@{}}Approximate \\ Optimisation\\ (SOCP, VQE, QAOA) \\
-    \cite{Mazumdar2020}, \cite{Venturelli2019}, \cite{Phillipson2020}, \cite{Rebentrost2018a}  \end{tabular} &
+    \cite{Mazumdar2020}, \cite{Buonaiuto2023}, \cite{Phillipson2020}, \cite{Venturelli2019}  \end{tabular} &
   \begin{tabular}[c]{@{}c@{}}Gate-based quantum computer \\ Quantum Annealer\end{tabular} \\ \hline  
@@
   \begin{tabular}[c]{@{}c@{}}Search and Count\\ (QAE, QPA, QPE) \\ 
-  \cite{Nakaji2020},\cite{Woerner2019}, \cite{Stamatopoulos2020}, \cite{Rebentrost2022} \cite{Rebentrost2018} \end{tabular} &
+  \cite{Woerner2019},\cite{Nakaji2020}, \cite{Stamatopoulos2020}, \cite{Rebentrost2018}, \cite{Rebentrost2022} \end{tabular} &
 \begin{tabular}[c]{@{}c@{}} Gate-based quantum \\ computer\end{tabular} \\ \hline \hline
@@
 \end{table*}
+
+expected to provide a quadratic speedup over the classical Monte Carlo methods used for this task.
 
@@
 
-Although this work is focused on two financial applications, the quantum algorithms reviewed here can be applied to a wide range of financial problems.
-In recent years, important surveys have been published on the topic of quantum computing in finance \cite{Orus2019,Egger2020a,Albareti2022,Herman2022,Herman2023}
-and we refer the interested readers to these works for a more comprehensive overview of the field.
+It is worth mentioning that important surveys have been published on the topic of quantum computing in finance in recent years.
+Among these, we highlight the works of Orús et al.
+\cite{Orus2019}, Egger et al.
+\cite{Egger2020a}, Albareti et al.
+\cite{Albareti2022}, Herman et al.
+\cite{Herman2022}, and Herman et al.
+\cite{Herman2023}, which have been instrumental in the development of this work.
+We refer the interested readers to these works for a more comprehensive overview of the field.
 
 \section{background}\label{sec:background}
+\subsection{Some basics concepts}
+\subsubsection{Assets and Portfolio Optimisation}
 
-\subsection{Some basics concepts}
+In finance, an asset is a financial instrument with economic value.
+It can be owned by an individual, a corporation, or a country with the expectation that it will provide a future benefit \cite{zotero-1119}.
+Financial assets can include stocks, bonds, commodities, derivatives, etc.
 
-\subsubsection{Assets and Portfolio Optimisation}
+On the other hand, a portfolio is a collection of individual investments.
+These investments can include a wide range of asset classes.
+Some key components of a portfolio include diversification, risk and return, and asset allocation.
+The goal of a portfolio is to diversify investments to achieve some specific objective such as maximising returns, minimising risk, or a balance between them \cite{Wilmott2007}.
+In order to achieve this, portfolio optimisation is used to find the best allocation of assets that fulfils the investor's objectives.
+
+In the original work by Markowitz \cite{Markowitz1952}, a mean-variance portfolio optimisation (MVPO) is proposed, which is widely 
+used used in the financial industry to this day, and whose formulation is presented in this section.
+
+The input for the MVPO is data about the historical prices and return of certain financial assets. An asset range $(1 \leq i \leq N)$ is considered, and a time range $(0 \leq t \leq T)$.
+
+Then, the first information we extract from the data is a list $P$ of current prices $P_i$  of the considered assets.
+
+\begin{equation}
+  \label{eq:1}
+  P_i = p_i(T)
+\end{equation}
+
+In addition, the historical returns of the assets between time $t-1$ and $t$ are calculated as:
+
+\begin{equation}
+  \label{eq:2}
+  r_i(t) = \frac{p_i(t) - p_i(t-1)}{p_i(t-1)}
+\end{equation}
+
+But for inference purposes, we must consider the expected return of the assets. So, given the historical returns, we can calculate the expected return of the assets as:
+
+\begin{equation}
+  \label{eq:3}
+  \mu_i = \frac{1}{T} \sum_{t=1}^{T} r_i(t)
+\end{equation}
+
+The variance of the returns of each asset and the covariance between returns of different assets over the historical period are given by:
+
+\begin{equation}
+  \label{eq:4}
+  \sigma_i^2 = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)^2
+\end{equation}
+
+\begin{equation}
+  \label{eq:5}
+  \sigma_{ij} = \frac{1}{T-1} \sum_{t=1}^{T} (r_i(t) - \mu_i)(r_j(t) - \mu_j)
+\end{equation}
 

### 2024-05-29 project/paper-v2.tex
update project paper

@@
 Then, the first information we extract from the data is a list $P$ of current prices $P_i$  of the considered assets.
-
 \begin{equation}
@@
 \end{equation}
-
 In addition, the historical returns of the assets between time $t-1$ and $t$ are calculated as:
-
 \begin{equation}
@@
 \end{equation}
-
 But for inference purposes, we must consider the expected return of the assets. So, given the historical returns, we can calculate the expected return of the assets as:
-
 \begin{equation}
@@
 \end{equation}
-
 The variance of the returns of each asset and the covariance between returns of different assets over the historical period are given by:
-
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
+A set of investments $x_i$ (given as a fraction of the total budget or number of asset units) allocated for each $i$th considered asset, 
+is then defined. Therefore, an optimal method for asset allocation aims to maximise the portfolio return 
+$\mu^\text{T} x$ while minimising risk, defined as the portfolio variance $x^\text{T}\Sigma x$, where $\mu$ is the vector of expected returns for each asset $i$ computed by Eq.
+\ref{eq:3}, $\Sigma$ is the covariance matrix calculate byt Eq.
+\ref{eq:4} and Eq.
+\ref{eq:5}, and $x$ is the vector
+of asset allocations measured as fractions of the total budget.
+Therefore, the problem of finding the optimal portfolio is reduced to find the $x$ vector that maximises the following objective function:
+\begin{equation}
+  \label{eq:6}
+  \mathcal{L}(x) = \mu^\text{T} x - q x^\text{T}\Sigma x
+\end{equation}
+where $q$ is a parameter that determines the trade-off between risk and return.  In a realistic scenario, budget constraints and non-negative 
+$x_i$ values (only buying) must be considered. Therefore, in the general case, the problem is formulated as follows:
+\begin{equation}
+\begin{aligned}
+  \label{eq:7}
+  \max_{x} \mathcal{L}(x): & \max_{x}\left(\mu^\text{T} x - q x^\text{T}\Sigma x \right) \\ 
+  \text{s.t.} & \quad \sum_{i=1}^{N} x_i = 1 
+\end{aligned}
+\end{equation}
+In section \ref{sec:literature1} we will review how this problem can be mapped into a QUBO problem and solved using quantum algorithms.
 
 \subsubsection{Option Pricing}
+
+An option is finan
 

### 2024-06-03 project/paper-v2.tex
update project paper sec 2

@@
 In Sec.
-\ref{sec:background} we introduce necessary concepts in finance and quantum computing.
+\ref{sec:background} we introduce necessary concepts in finance and quantum algorithms.
 Then, in Sec.
@@
 
-In finance, an asset is a financial instrument with economic value.
-It can be owned by an individual, a corporation, or a country with the expectation that it will provide a future benefit \cite{zotero-1119}.
+In finance, an asset is a financial instrument that can be owned by an individual, a corporation, or a country with the expectation that it will provide a future benefit \cite{2023}.
 Financial assets can include stocks, bonds, commodities, derivatives, etc.
 
-On the other hand, a portfolio is a collection of individual investments.
-These investments can include a wide range of asset classes.
+On the other hand, a portfolio is a collection of individual investment that can include a wide range of asset classes.
 Some key components of a portfolio include diversification, risk and return, and asset allocation.
@@
 
-The input for the MVPO is data about the historical prices and return of certain financial assets. An asset range $(1 \leq i \leq N)$ is considered, and a time range $(0 \leq t \leq T)$.
+The input for the MVPO is data about the historical prices and returns of certain financial assets. An asset range $(1 \leq i \leq N)$ is considered, and a time range $(0 \leq t \leq T)$.
 
@@
 $\mu^\text{T} x$ while minimising risk, defined as the portfolio variance $x^\text{T}\Sigma x$, where $\mu$ is the vector of expected returns for each asset $i$ computed by Eq.
-\ref{eq:3}, $\Sigma$ is the covariance matrix calculate byt Eq.
+\ref{eq:3}, $\Sigma$ is the covariance matrix calculate by Eq.
 \ref{eq:4} and Eq.
@@
 \subsubsection{Option Pricing}
+An option is a financial instrument that gives the holder the opportunity but not the obligation to buy or sell the  \textbf{underlying asset} at an agreed-upon price (\textbf{strike price}) at a specified time in the future (\textbf{expiration date}) \cite{Wilmott2007}.
+There are many kind of options,
+but the most common are \textbf{call} and \textbf{put options} (buy and sell options, respectively). Furthermore, we must 
+differentiate between European and American options.
+The former can only be exercised at the expiration date, while the latter can be exercised at any time before the expiration date \cite{Rebentrost2018}.
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{pay-off.pdf}
+  \caption{\label{fig:profit-diagram} Profit diagram for a call option. Adapted from \cite{Wilmott2007}.}
+\end{figure}
 
-An option is finan
+The profit diagram for a call option is shown in Fig. \ref{fig:profit-diagram}, where the \textbf{premium} is the price paid for contract initially. Then, the 
+intrinsic value of the option at the expiration date is calculated as the difference between the stock price $S$ and the strike price $E$. 
+This function of the underlying asset is called the \textbf{payoff function} and is given by:
+\begin{equation}
+  \label{eq:8}
+  \text{Payoff} = \max(S - E, 0)
+\end{equation}
+The goal of option pricing is to estimate the option payoff at the expiration date and then discount that value to determine the \textbf{fair value}, which indicates the amount of money an 
+investor should pay to enter the option contract today, given the current market conditions. 
+The challenging part is to compute the expected value of the option payoff, which is usually done using Monte Carlo methods.
+In section \ref{sec:literature2} we will review how this problem can be solved using quantum algorithms.
 
-\subsubsection{Quantum Computing}
+\subsubsection{Variational Quantum Algorithms}
+
+Variational Quantum Algorithms (VQAs) are hybrid quantum-classical algorithms that 
+laverage some quantum mechanical principles to solve optimisation problems.
+The main goal of a VQA is to find an approximate solution for optimisation and combinatorial problems \cite{Cerezo2021}.
+Having stated the optimisation problem,
+we must map its cost function $C(\vec{\theta})$ to a Hamiltonian operator, whose ground state energy encodes the optimal solution.
+
+\begin{figure}[h!]
+\centering 
+\includegraphics[width=0.45\textwidth]{VQE-Egger.pdf}
+  \caption{\label{fig:vqe} Variational Quantum Eigensolver (VQE) algorithm.}
+\end{figure}
+
+One of this algorithms is the Variational Quantum Eigensolver
+
+\begin{figure*}[ht!]
+  \begin{center} 
+  \includegraphics[width=\textwidth]{QAOA-Blekos.pdf}
+  \caption{\label{fig:qaoa} Quantum Approximate Optimisation Algorithm (QAOA) algorithm.}
+  \end{center}
+\end{figure*}
+
+\subsubsection{Quantum Amplitude Estimation}
 

### 2024-06-03 project/paper-v2.tex
update project paper VQE

@@
 \usepackage{hhline}
-\bibliographystyle{apsrev4-2}
+%\bibliographystyle{apsrev4-2}
 \def\bibsection{\section{\refname}} 
@@
 
-On the other hand, a portfolio is a collection of individual investment that can include a wide range of asset classes.
-Some key components of a portfolio include diversification, risk and return, and asset allocation.
-The goal of a portfolio is to diversify investments to achieve some specific objective such as maximising returns, minimising risk, or a balance between them \cite{Wilmott2007}.
+On the other hand, a portfolio is a set of individual investments that may include a wide range of asset classes.
+A portfolio is characterised by components such as diversification, risk and return, and asset allocation.
+The main goal of a portfolio is to diversify investments to achieve some specific objective such as maximising returns, minimising risk, or a balance between them \cite{Wilmott2007}.
 In order to achieve this, portfolio optimisation is used to find the best allocation of assets that fulfils the investor's objectives.
 
-In the original work by Markowitz \cite{Markowitz1952}, a mean-variance portfolio optimisation (MVPO) is proposed, which is widely 
-used used in the financial industry to this day, and whose formulation is presented in this section.
+In the original work by Markowitz \cite{Markowitz1952}, a mean-variance portfolio optimisation (MVPO) is proposed, which is widely used used in the financial industry to this day, and whose formulation is presented in this section.
 
-The input for the MVPO is data about the historical prices and returns of certain financial assets. An asset range $(1 \leq i \leq N)$ is considered, and a time range $(0 \leq t \leq T)$.
+The input for the MVPO is the historical prices and returns of the considered financial assets. An asset range $(1 \leq i \leq N)$ is considered, and a time range $(0 \leq t \leq T)$.
 
-Then, the first information we extract from the data is a list $P$ of current prices $P_i$  of the considered assets.
+Then, the first step is to obtain a list $P$ of current prices $P_i$  of the assets of interest.
 \begin{equation}
@@
 \end{equation}
-The goal of option pricing is to estimate the option payoff at the expiration date and then discount that value to determine the \textbf{fair value}, which indicates the amount of money an 
+The purpose of option pricing is to estimate the option payoff at the expiration date and then, determine the \textbf{fair value}, which indicates the amount of money an 
 investor should pay to enter the option contract today, given the current market conditions. 
@@
 Variational Quantum Algorithms (VQAs) are hybrid quantum-classical algorithms that 
-laverage some quantum mechanical principles to solve optimisation problems.
-The main goal of a VQA is to find an approximate solution for optimisation and combinatorial problems \cite{Cerezo2021}.
-Having stated the optimisation problem,
-we must map its cost function $C(\vec{\theta})$ to a Hamiltonian operator, whose ground state energy encodes the optimal solution.
+are well suited for NISQ devices and allow us to get approximate solutions to optimisation and combinatorial problems efficiently \cite{Cerezo2021}. This algorithms rely on the fact that we can, in 
+principle, reformulate any optimisation problem as a QUBO problem \cite{Glover2019}.
 
+QUBO problems can not directly be cast into a quantum computer, but they share a isomorphism with the Ising Hamiltonian, making it easy to map a QUBO problem into an Ising Hamiltonian, which can be cast into adequate quantum operators in a quantum computer \cite{Buonaiuto2023}.
 \begin{figure}[h!]
 \centering 
-\includegraphics[width=0.45\textwidth]{VQE-Egger.pdf}
-  \caption{\label{fig:vqe} Variational Quantum Eigensolver (VQE) algorithm.}
+\includegraphics[width=0.5\textwidth]{VQE-Buonaiuto.pdf}
+  \caption{\label{fig:vqe} Variational Quantum Eigensolver (VQE) algorithm. Adapted from \cite{Buonaiuto2023}.}
 \end{figure}
-
-One of this algorithms is the Variational Quantum Eigensolver
 

### 2024-06-04 project/paper-v2.tex
update paper

@@
 \end{equation}
+
 In addition, the historical returns of the assets between time $t-1$ and $t$ are calculated as:
@@
 \end{equation}
+
 But for inference purposes, we must consider the expected return of the assets. So, given the historical returns, we can calculate the expected return of the assets as:
@@
 \end{equation}
+
 The variance of the returns of each asset and the covariance between returns of different assets over the historical period are given by:
@@
 \end{equation}
+
 A set of investments $x_i$ (given as a fraction of the total budget or number of asset units) allocated for each $i$th considered asset, 
@@
 \end{equation}
-The purpose of option pricing is to estimate the option payoff at the expiration date and then, determine the \textbf{fair value}, which indicates the amount of money an 
+
+The purpose of option pricing is to estimate Eq. \ref{eq:8} at the expiration date and then, determine the \textbf{fair value}, which indicates the amount of money an 
 investor should pay to enter the option contract today, given the current market conditions. 
@@
 \subsubsection{Variational Quantum Algorithms}
-
 Variational Quantum Algorithms (VQAs) are hybrid quantum-classical algorithms that 
@@
 
+One of the VQAs used for portfolio optimisation is the Variational Quantum Eigensolver (VQE) algorithm (see Fig.
+\ref{fig:vqe}).
+The aim of the VQE algorithm is to find an approximate value for the ground state energy of a given Hamiltonian.
+
+The starting point is to define a cost function
+\begin{equation}
+  \label{eq:9}
+  C(\boldsymbol{\theta}) =\langle\psi(\boldsymbol{\theta})|H|\psi(\boldsymbol{\theta})\rangle
+\end{equation}
+where $H$ is the Hamiltonian of our problem, $\boldsymbol{\theta}$ is a set of trainable parameters, and a trial state $|\psi(\boldsymbol{\theta})\rangle = U(\boldsymbol{\theta}) |0\rangle ^{\otimes n}$ for some ansatz $U(\boldsymbol{\theta})$
+
 \begin{figure*}[ht!]
   \begin{center} 
-  \includegraphics[width=\textwidth]{QAOA-Blekos.pdf}
-  \caption{\label{fig:qaoa} Quantum Approximate Optimisation Algorithm (QAOA) algorithm.}
+  \includegraphics[width=0.85\textwidth]{QAOA-Blekos.pdf}
+  \caption{\label{fig:qaoa} Quantum Approximate Optimisation Algorithm (QAOA) algorithm. Adapted from \cite{Blekos2024}.}
   \end{center}

### 2024-06-04 project/paper-v2.tex
update project paper VQE

@@
 
-\title{Quantum Algorithms for Portflolio Optimisation in Finance}
+\title{Quantum Algorithms for Portflolio Optimisation and Option pricing}
 
@@
 investor should pay to enter the option contract today, given the current market conditions. 
-The challenging part is to compute the expected value of the option payoff, which is usually done using Monte Carlo methods.
-In section \ref{sec:literature2} we will review how this problem can be solved using quantum algorithms.
+The challenging part is to compute the expected value of the option payoff, which is usually done using Monte Carlo methods. In section \ref{sec:literature2} we will discuss the quantum approach to 
+this problem.
 
-\subsubsection{Variational Quantum Algorithms}
+\subsection{Variational Quantum Algorithms}
 Variational Quantum Algorithms (VQAs) are hybrid quantum-classical algorithms that 
@@
 \end{equation}
-where $H$ is the Hamiltonian of our problem, $\boldsymbol{\theta}$ is a set of trainable parameters, and a trial state $|\psi(\boldsymbol{\theta})\rangle = U(\boldsymbol{\theta}) |0\rangle ^{\otimes n}$ for some ansatz $U(\boldsymbol{\theta})$
-
+where $H$ is the Hamiltonian of our problem, that can be written as the weighted sum of Pauli operators \cite{Tilly2022}, such
 \begin{figure*}[ht!]
@@
   \includegraphics[width=0.85\textwidth]{QAOA-Blekos.pdf}
-  \caption{\label{fig:qaoa} Quantum Approximate Optimisation Algorithm (QAOA) algorithm. Adapted from \cite{Blekos2024}.}
+  \caption{\label{fig:qaoa} Quantum Approximate Optimisation Algorithm (QAOA). Adapted from \cite{Blekos2024}.}
   \end{center}
 \end{figure*}
+that: 
+\begin{equation}
+  \label{eq:10}
+  \mathcal{H} = \sum_{i} c_i P_i
+\end{equation}
+where $P_i \in \{I, X, Y, Z\}^{\otimes N}$ are called Pauli strings and $N$ is the number of qubits, and $c_i$ are the coefficients of the Pauli strings.
+Additionally, $\boldsymbol{\theta}$ is a set of trainable parameters for some trial state $|\psi(\boldsymbol{\theta})\rangle = U(\boldsymbol{\theta}) |0\rangle ^{\otimes n}$.
+So, Eq.
+\ref{eq:9} becomes:
+\begin{equation}
+  \label{eq:11}
+  C(\boldsymbol{\theta}) =\sum_i c_i \langle \boldsymbol{0}|U(\boldsymbol{\theta})^\dagger P_i U(\boldsymbol{\theta})|\boldsymbol{0}\rangle
+\end{equation}
+for some ansatz $U(\boldsymbol{\theta})$, that can be written as
+\begin{equation}
+  \label{eq:12}
+  U(\boldsymbol{\theta}) = \prod_{k=1}^{p} e^{-i\theta_k \mathcal{H}_k}
+\end{equation}
+where $\mathcal{H}_k$ are the terms of the Hamiltonian $\mathcal{H}$  and $p$ determines the precision of the approximation \cite{Cerezo2021}.
+
+Therefore, the VQE consists of classical algorithm that uses a quantum algorithm as a subroutine. We first initialise the state 
+$|\psi(\boldsymbol{\theta})\rangle$ with some initial parameters $\boldsymbol{\theta}$, then we measure Eq. \ref{eq:11} and use a classical optimisation algorithm to update 
+the parameters $\boldsymbol{\theta}$ until we reach a minimum value of $C(\boldsymbol{\theta})$.
+
+The variational principle ensures that at some point the VQE algorithm will converge to:
+\begin{equation}
+  \label{eq:13}
+  E_0 \leq \frac{\langle \psi(\boldsymbol{\theta})|\mathcal{H}|\psi(\boldsymbol{\theta})\rangle}{\langle \psi(\boldsymbol{\theta})|\psi(\boldsymbol{\theta})\rangle}
+\end{equation}
+where $E_0$ is the ground state energy of the Hamiltonian $\mathcal{H}$ and $|\psi(\boldsymbol{\theta})\rangle$ is some trial state.
+
+In a similar fashion, the Quantum Approximate Optimisation Algorithm (QAOA) is another VQA that was first 
+conceived as
 

### 2024-06-05 project/paper-v2.tex
update project paper QAOA

@@
 In a similar fashion, the Quantum Approximate Optimisation Algorithm (QAOA) is another VQA that was first 
-conceived as
+conceived by Farhi et al. \cite{Farhi2014} for combinatorial optimisation problems (e.g. Max-Cut).
+
+Mathematically, we can see the combinatorial optimisation problem as a graph with $m$ edges and $n$ vertices. Then, define a bitstring $x = $
+
+\begin{figure}[h!]
+\centering 
+\includegraphics[width=0.35\textwidth]{combinatorial-problem.pdf}
+  \caption{\label{fig:combinatorial} A graph representation of a combinatorial optimisation problem. Adapted from \cite{Baker2022}.} 
+\end{figure}
 

### 2024-06-05 project/paper-v2.tex
update project paper QAOA

--- 2024-06-04 project/paper-v2.tex
@@
 In a similar fashion, the Quantum Approximate Optimisation Algorithm (QAOA) is another VQA that was first 
-conceived as
+proposed by Farhi et al. \cite{Farhi2014} for combinatorial optimisation problems (e.g. Max-Cut).
 
-\subsubsection{Quantum Amplitude Estimation}
+Mathematically, we can see the combinatorial optimisation problem as a graph with $m$ edges and $n$ vertices (see Fig.
+\ref{fig:combinatorial}).
+Then, define a binary string $x = \{x_1,\cdots, x_n\}$ with $x_i \in \{0, 1\}$.
+\begin{figure}[h!]
+\centering 
+\includegraphics[width=0.35\textwidth]{combinatorial-problem.pdf}
+  \caption{\label{fig:combinatorial} A graph representation of a combinatorial optimisation problem with $n =4$ vertices and $m=6$ edges. Adapted from \cite{Baker2022}.} 
+\end{figure}
+The goal is to find the binary string that maximises the objective function: 
+\begin{equation}
+  \label{eq:14}
+  C(x) = \sum_{i,j=1}^{n} \Sigma_{i,j} x_i (1-x_j)
+\end{equation}
+
+In Fig. \ref{fig:qaoa} we show the general structure of the QAOA. First, it initialise $n$ states as a uniform superposition
+\begin{equation}
+  \label{eq:15}
+  |+\rangle^{\otimes n} = \frac{1}{\sqrt{2^n}} \sum_{x} |x\rangle 
+\end{equation}
+and then, it applies a sequence of $p$ anzates $U(\boldsymbol{\gamma}, \boldsymbol{\beta})$  that are defined as 
+\begin{equation}
+  \label{eq:16}
+  U(\boldsymbol{\gamma}, \boldsymbol{\beta}) = \prod_{k=1}^{p} e^{-i\beta_k \mathcal{H}_M} e^{-i\gamma_k \mathcal{H}_C}
+\end{equation}
+where $\mathcal{H}_M$ and $\mathcal{H}_C$ are the mixing and cost Hamiltonians, respectively.
+The cost Hamiltonian $\mathcal{H}_C$ encodes the cost function of the problem and the mixing Hamiltonian helps guide the optimisation in Hilbert space towards ground state of $\mathcal{H}_C$ \cite{Moll2018}.
+Similar to the VQE, the cost Hamiltonian is defined by Eq.
+\ref{eq:10} and the mixer Hamiltonian
+is defined as $\mathcal{H}_M = -\sum_{i} \sigma_i^x$ \cite{Moll2018}.
+
+After applying the anzates $p$ times, we measure the expectation value of the cost Hamiltonian $C(\boldsymbol{\gamma},\boldsymbol{\beta}) = \langle \psi(\boldsymbol{\gamma}, \boldsymbol{\beta})|\mathcal{H}_C|\psi(\boldsymbol{\gamma},\boldsymbol{\beta})\rangle$ and use a classical optimisation algorithm to update the parameters $\boldsymbol{\gamma}$ and $\boldsymbol{\beta}$ until we reach a minimum value of $C(\boldsymbol{\gamma},\boldsymbol{\beta})$.
+
+\subsection{Quantum Amplitude Estimation}
 

### 2024-06-07 project/paper-v2.tex
update paper

@@
 investor \cite{Herman2022}.
-Machine Learning based models (ML) such as the one proposed by Mazumdar et al.
-\cite{Mazumdar2020} are widely used for minimising risk contributors and maximising asset diversification.
+Monte Carlo based models (MC) such as the one proposed by Shadabfar and Cheng \cite{Shadabfar2020} are widely used for minimising risk contributors and maximising asset diversification.
 
 In particular, quantum approaches to portfolio optimisation are gaining significant attention \cite{Quarterly2020}.
-In 2019 Kerenidis, Prakash and Szilágyi proposed a quantum algorithm for the constrained portfolio optimisation problem.
+In 2019 Kerenidis, Prakash and Szilágyi \cite{Kerenidis2019},  proposed a quantum algorithm for the constrained portfolio optimisation problem.
 Based on an interior point method, this algorithm reduces the optimisation problem to a second order cone program (SOCP).
@@
 where $\mathcal{H}_M$ and $\mathcal{H}_C$ are the mixing and cost Hamiltonians, respectively.
-The cost Hamiltonian $\mathcal{H}_C$ encodes the cost function of the problem and the mixing Hamiltonian helps guide the optimisation in Hilbert space towards ground state of $\mathcal{H}_C$ \cite{Moll2018}.
-Similar to the VQE, the cost Hamiltonian is defined by Eq.
+$\mathcal{H}_C$ encodes the cost function of the problem and $\mathcal{H}_M$ helps guide the optimisation in Hilbert space towards ground state of $\mathcal{H}_C$ \cite{Moll2018}.
+Similar to the VQE, $\mathcal{H}_C$ can de defined by Eq.
 \ref{eq:10} and the mixer Hamiltonian
@@
 
-\section{Portfolio Optimisation}\label{sec:literature1}
+Quantum Amplitude Estimation (QAE) is an algorithm that can estimate a parameter $a$ with a convergence rate of $\mathcal{O}(1/M)$, where $M$ is the number of samples, which corresponds to a quadratic speedup compared to classical methods (i.e Monte Carlo) which requires $\mathcal{O}(1/\sqrt{M})$ \cite{Egger2020}.
+QAE is based on a unitary operator $\mathcal{A}$  that acts on $(n+1)$ qubits such that
+\begin{equation}
+  \label{eq:17}
+  \mathcal{A}|0\rangle_{n+1} = \sqrt{1-a}|\psi_0\rangle_n|0\rangle + \sqrt{a}|\psi_1\rangle_n|1\rangle
+\end{equation}
+for some normalised states $|\psi_0\rangle_n$ and $|\psi_1\rangle_n$, where $a \in [0,1]$ is unknown \cite{Stamatopoulos2020}. We can then, 
+infer the value of $a$ by performing multiple measurements on $|\Psi\rangle$ and obtaining the ratio of the good state.
+This method does not represent any advantage with respect to the classical method since the number of queries is exactly the same.
 
-\subsection{Classical Portfolio Optimisation}
+Therefore, QAE is used along with the amplitude amplification algorithm to amplify the probability of measuring the good state. We achieve that by applying the following operator: 
+\begin{equation}
+  \label{eq:18}
+  \mathcal{Q} = -\mathcal{A} \mathcal{S}_0 \mathcal{A} \mathcal{S}_{\chi}
+\end{equation}
+where $\mathcal{S}_0$ is the inversion about the mean operator and $\mathcal{S}_{\chi}$ is the oracle operator that flips the sign of the good state \cite{Suzuki2020}.
 
-\subsection{Quantum Portfolio Optimisation}
+By defining $\theta_a \in [0, 2\pi]$  so that $a = \sin^2(\theta_a)$, and applying repeatedly $\mathcal{Q}$ for $m$ times on $|\Psi\rangle$  results in
+\begin{equation}
+  \label{eq:19}
+\begin{aligned}
+  \mathcal{Q}^m |\Psi\rangle  &= \sin((2m + 1)\theta_a)|\psi_1\rangle_n|1\rangle\\
+  & + \cos((2m + 1)\theta_a)|\psi_0\rangle_n|0\rangle 
+\end{aligned}
+\end{equation}
+once measured we can apply a maximum likelihood estimation to infer the value of $a$ \cite{Suzuki2020}.
 
-\section{Option Pricing}\label{sec:literature2}
+\section{Classical Approaches}\label{sec:literature1}
 
-\subsection{Classical Option Pricing}
+In this section we review the main classical approaches to portfolio optimisation and option pricing. 
+This section is not intended to be an detailed review on the topic, but rather to provide a point of comparison with the quantum approaches in Sec. \ref{sec:literature2}
 
-\subsection{Quantum Option Pricing}
+\subsection{Portfolio Optimisation}
+
+Classical approaches for portfolio optimisation are based on the MVPO (discussed in Sec. \ref{sec:background}).  
+Shadabfar and  Cheng \cite{Shadabfar2020} proposed a probabilistic method for portfolio optimisation using a hybrid Monte Carlo simulation and Markowitz model.
+They use the cost function defined by Eq.
+\ref{eq:6}, and apply a Monte Carlo to generate multiple random samples of the asset allocations.
+Then, using the Markowitz optimisation model they computed the optimum values of return and risk for each portfolio.
+\begin{figure}[h!]
+\centering 
+\includegraphics[width=0.5\textwidth]{return-risk-error.pdf}
+  \caption{\label{fig:monte-carlo} Convergence of the probabilistic method proposed by Shadabfar and Cheng.
+The convergence of the return and risk probability depends on the number of random samples considered.
+Adapted from \cite{Shadabfar2020}.}
+\end{figure}
+
+As shown in Fig.
+\ref{fig:monte-carlo}, they found that the convergence of their method is dependent on the number of samples used in the Monte Carlo method.
+Additionally, they demonstrated that the model results
+are roubust to errors in the model's input data, indicating the stability and reliability of their method.
+
+\subsection{Option Pricing}
+
+ The most common classical approach to option pricing are the Monte Carlo methods. In contrast to the Black-Scholes model, which can provide 
+ analytical solutions for very simplistic cases, Monte Carlo methods due to their random nature can model and capture the 
+ complex behaviour of the financial markets.
+
+Stochastic differential equations (SDEs) are used to model the evolution of the underlying asset price.
+The most common SDE used for this task is the Geometric Brownian Motion (GBM) model, which is given by:
+
+\begin{equation}
+  \label{eq:20}
+  S_{t+1} = S_t \exp\left(\left[\mu - \frac{1}{2}\sigma^2\right]\Delta t + \sigma \sqrt{\Delta t}Z_t\right)
+\end{equation}
+
+\section{Quantum Approaches}\label{sec:literature2}
+
+\subsection{Portfolio Optimisation}
+
+\subsection{Option Pricing}
 

### 2024-06-10 project/paper-v2.tex
Final version paper

@@
 Stochastic differential equations (SDEs) are used to model the evolution of the underlying asset price.
-The most common SDE used for this task is the Geometric Brownian Motion (GBM) model, which is given by:
-
+The most common SDE used for this task is the Geometric Brownian Motion (GBM) model \cite{Sood2019}, which is given by:
 \begin{equation}
@@
 \end{equation}
+where $S_t$ is the asset price at time $t$, $\mu$ is the expected return, $\sigma$ is the volatility, $\Delta t$ is the time step, and $Z_t$ is a standard normal random variable.
+
+The general idea of this method is to simulate the asset price evolution multiple times and calculate the payoff of the option at the expiration date (See Fig.
+\ref{fig:monte-carlo-option}).
+The expected value of the option price is then calculated as the average of the payoffs.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{monte-carlo-paths.pdf}
+  \caption{\label{fig:monte-carlo-option} Monte Carlo simulation of the option price for a maturity time $T =1$ (years).
+The option price is calculated as the average of the payoffs of the option at the expiration date.
+Adapted from \cite{Sood2019}.}
+\end{figure}
+
+In general, the estimation error goes as $\mathcal{O}(1/\sqrt{M})$, where $M$ is the number of samples used in the Monte Carlo simulation.
+Therefore, we need to find a balance between the estimation error and the time required to compute all the random samples.
+Although a simplified version of the Monte Carlo method is provided in this section,
+this is the basis for more complex methods.
 
 \section{Quantum Approaches}\label{sec:literature2}
+Having reviewed the classical approaches for both portfolio optimisation and option pricing, it is time to 
+introduce the quantum approaches to these problems. In this section we will review the implementations of the quantum algorithms discussed in Sec. \ref{sec:background} for these tasks.
+\subsection{Portfolio Optimisation}
+In 2023, Buonaiuto et al.
+\cite{Buonaiuto2023} published a paper where they implemented the VQE algorithm in multiple simulations and in real quantum computers.
+They used four assets (Apple, IBM, Netflix and Tesla)
+historical data, and the MVPO model to state the optimisation problem. 
+\begin{figure}[h!]
+\centering 
+\includegraphics[width=0.45\textwidth]{data-paper.png}
+  \caption{\label{fig:data-set} Sample of the historical data of of the assets Apple, IBM, Netflix and Tesla used for the VQE implementation by 
+Biunaiuto et al. \cite{Buonaiuto2023}.}
+\end{figure}
 
-\subsection{Portfolio Optimisation}
+They mapped the MVPO problem into a QUBO problem (Eq. \ref{eq:21} and Eq. \ref{eq:22}), and then converted it into quantum operators using 
+the Ising Hamiltonian, as proposed by \cite{Tilly2022}.  Prices, returns, and covariance matrix of the assets were calculated using the historical data (see Fig. \ref{fig:data-set}). 
+\begin{equation}
+  \label{eq:21}
+  \max_{b} \mathcal{L}(b) =\max_{b}\left(\mu''^{T} b - qb^{T} \Sigma'' b - \lambda(P''^{T}b- 1)^2\right)
+\end{equation}
+\begin{equation}
+  \label{eq:22}
+  \mathcal{H} = \sum_{i} h_i Z_i + \sum_{i,j} J_{ij} Z_i\otimes Z_j + \lambda \left(\sum_{i}\pi_i Z_i -\beta\right)^2
+\end{equation}
+
+The total expendable budget was set to $B=2000$ for all the experiments, a risk aversion parameter $q=0.5$ was considered.
+The initial ansatz parameters were randomly selected between $[-\pi, \pi]$.
+A number of 200 epochs was set, and a penalty parameter $\lambda=10$  was chosen.
+Finally, the Coybla classical optimiser was used.
+
+The results of the VQE implementation are shown in Fig.
+\ref{fig:results}.
+The results were contrasted with a classical solution.
+The Toronto, Kolkata and Auckland quantum computers performances perfectly match the classical result.
+Different ansatz were used for the different quantum computers.
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=.90\textwidth]{results-vqe.png}
+  \caption{\label{fig:results} Results of the VQE implementation by Buonaiuto et al. \cite{Buonaiuto2023}. The optimal asset allocation for the four assets considered is shown. Extracted from 
+  \cite{Buonaiuto2023}}
+\end{figure*}
 
 \subsection{Option Pricing}
+In 2020, Stamatopoulos et al.
+\cite{Stamatopoulos2020} published a work where they presented a methodology to price options and portfolios of options on a gate-based quantum computer using the QAE algorithm.
+They implemented their methodology using the IBM Q Tokyo quantum device.
+Among
+the options covered in their work, they considered multi-asset options and path dependent options.
 
-\section{Hardware Implementations}\label{sec:hardware}
+They discussed the QAE implementation along with the quantum Fourier transform.
+In their work, they prove mathematically the potential speedup the QAE algorithm could provide compared to the classical Monte Carlo methods, which is in agreement with \cite{Woerner2019, Nakaji2020,Rebentrost2018, Rebentrost2022}.
+All these works suggest that the estimation error of QAE goes as $\mathcal{O}(1/M)$, whereas the classical Monte Carlo method provides an estimation error that scales as $\mathcal{O}(1/\sqrt{M})$, where $M$ is the number of samples.
+
+Instead of using the QFT, the approach proposed by Suzuki et al. \cite{Suzuki2020} and discussed in Sec. \ref{sec:background} was considered.
+Results of the QAE implementation are shown in Fig. \ref{fig:results-qae}. The authors point out the good performance of this algorithm compared to the classical MC.
+
+Furthermore, they used their methodology to price a portfolio of options. They used these results to experimentally prove the speedup of the QAE algorithm (see Fig. \ref{fig:portfolio-option}).
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.45\textwidth]{option-price-estimation.pdf}
+  \caption{\label{fig:results-qae} Estimated option price using QAE.
+$S_0 = 2$, $ \sigma = 10\%$, $r = 4\%$ and $T=300/365$ were used.
+The red dotted line is the value computed using classical Monte Carlo, and the blue bars represent the estimated option price using QAE after 100000 samples.
+Adapted  from \cite{Stamatopoulos2020}.}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.45\textwidth]{estimation-error.pdf}
+  \caption{\label{fig:portfolio-option} 
+Convergence of the estimation error of the option price for a portfolio of options is shown.
+The yellow dotted line shows the error convergence for the QAE, whereas the blue dashed line indicates the error convergence for the classical Monte Carlo.
+Adapted  from \cite{Stamatopoulos2020}.}
+\end{figure}
 
 \section{Discussion and Conclusion}\label{sec:discussion}
+In this work we have reviewed the quantum algorithms proposed for financial applications.
+Algorithms such as the VQAs and QAE have demonstrated a potential application for portfolio optimisation and option pricing problems.
+
+VQAs have been successfully implemented for optimisation problems.
+The results show that the performance of these algorithms is highly dependent on the ansatz structure and hardware topology.
+Nevertheless, the performance of the VQAs is formidable even when no error mitigation protocol is considered, making them a good candidate for exploiting the computational power that NISQi devices offer to date.
+
+On the other hand, QAE has been experimentally proven to provide a quadratic speedup over the classical Monte Carlo methods. Applied to 
+option pricing, this algorithm has the potential to provide a significant benefit in terms of time and computational 
+resources.
+
+Finally, although the results of the quantum algorithms applied to financial applications are promising, 
+hardware limitations must be addressed as well as the development of error mitigation protocols. These  implementation results 
+show the potential of quantum algorithms beyond the theoretical framework, and pave the way for future research and 
+applications in the financial industry.
 

### 2024-06-11 project/paper-v2.tex
submit paper

@@
 We describe the ideas behind important quantum algorithms such as Variational Quantum Algorithms (VQAs) and Quantum Amplitude Estimation (QAE) and their theoretical performance.
-We provide an overview of the hardware available for their implementation and discuss the results
+We provide an overview of the hardware  implementations and discuss the results
 obtained by some study cases, comparing them with their classical counterparts. Finally, we discuss the future prospects of these algorithms in the financial industry.
@@
 
-In Sec.
-\ref{sec:hardware} we provide an overview of the hardware implementations of these algorithms.
-Finally, Sec.
-\ref{sec:discussion} concludes this work and discusses the future prospects of quantum algorithms in the financial industry.
+ Finally, Sec. \ref{sec:discussion} concludes this work and discusses the future prospects of quantum algorithms in the financial industry.
 
@@
 \end{equation}
-In section \ref{sec:literature1} we will review how this problem can be mapped into a QUBO problem and solved using quantum algorithms.
+Although analytical solutions exist for particular cases, the general problem is unsolvable efficiently with classical methods, and 
+heuristic methods such as Monte Carlo simulations are used to find approximate solutions \cite{Shadabfar2020}.
 
@@
 investor should pay to enter the option contract today, given the current market conditions. 
-The challenging part is to compute the expected value of the option payoff, which is usually done using Monte Carlo methods. In section \ref{sec:literature2} we will discuss the quantum approach to 
+The challenging part is to compute the expected value of the option payoff, which is usually done using Monte Carlo methods \cite{Sood2019}.
+In section \ref{sec:literature2} we will discuss the quantum approach to
 this problem.
@@
 Additionally, $\boldsymbol{\theta}$ is a set of trainable parameters for some trial state $|\psi(\boldsymbol{\theta})\rangle = U(\boldsymbol{\theta}) |0\rangle ^{\otimes n}$.
-So, Eq.
+Therefore, Eq.
 \ref{eq:9} becomes:
@@
 
-The variational principle ensures that at some point the VQE algorithm will converge to:
+The variational principle ensures that after enough iterations the VQE algorithm will converge to:
 \begin{equation}
@@
 
-Furthermore, they used their methodology to price a portfolio of options. They used these results to experimentally prove the speedup of the QAE algorithm (see Fig. \ref{fig:portfolio-option}).
+Furthermore, they used their methodology to price a portfolio of options, and using these results, they experimentally proved the speedup of the QAE algorithm (see Fig. \ref{fig:portfolio-option}).
 
@@
 The results show that the performance of these algorithms is highly dependent on the ansatz structure and hardware topology.
-Nevertheless, the performance of the VQAs is formidable even when no error mitigation protocol is considered, making them a good candidate for exploiting the computational power that NISQi devices offer to date.
+Nevertheless, the performance of the VQAs is formidable even when no error mitigation protocol is considered, making them a good candidate for exploiting the computational power that NISQ devices offer to date.
 
@@
 
-\bibliography{/home/pacman/Documents/git-repositories/quantum-computing/lecture-notes/bibliography.bib}
+\bibliography{./bibliography.bib}
 

### 2024-06-26 project/paper-v2.tex
corrections-figures

@@
 historical data, and the MVPO model to state the optimisation problem.
-\begin{figure}[h!]
+
+\begin{table}[h]
     \centering
-\includegraphics[width=0.45\textwidth]{data-paper.png}
-  \caption{\label{fig:data-set} Sample of the historical data of of the assets Apple, IBM, Netflix and Tesla used for the VQE implementation by 
-Biunaiuto et al. \cite{Buonaiuto2023}.}
-\end{figure}
+     \caption{Sample of the historical data of the assets Apple, IBM, Netflix and Tesla used for the VQE implementation by Biunaiuto et al. \cite{Buonaiuto2023}.}
+    \begin{tabular}{cccccc}
+        \hline
+        \textbf{N.} & \textbf{Date} & \textbf{AAPL} & \textbf{IBM} & \textbf{NFLX} & \textbf{TSLA} \\
+        \midrule
+        0  & 2011-12-23 & 12.260060 & 116.121811 & 10.374286 & 1.860000 \\
+        1  & 2011-12-27 & 12.357327 & 116.247513 & 10.085714 & 1.904667 \\
+        2  & 2011-12-28 & 12.239082 & 115.644180 & 9.885714 & 1.900667 \\
+        3  & 2011-12-29 & 12.314467 & 117.020638 & 9.900000 & 1.915333 \\
+        4  & 2011-12-30 & 12.310820 & 115.575020 & 9.898571 & 1.904000 \\
+        5  & 2012-01-03 & 12.500192 & 117.096062 & 10.320000 & 1.872000 \\
+        6  & 2012-01-04 & 12.567370 & 116.618378 & 11.492857 & 1.847333 \\
+        7  & 2012-01-05 & 12.706894 & 116.065247 & 11.328571 & 1.808000 \\
+        8  & 2012-01-06 & 12.839729 & 114.732758 & 12.327143 & 1.794000 \\
+        9  & 2012-01-09 & 12.819363 & 114.135643 & 14.025714 & 1.816667 \\
+        \hline
+    \end{tabular} 
+    \label{tab:stock_prices}
+\end{table}
 
 They mapped the MVPO problem into a QUBO problem (Eq. \ref{eq:21} and Eq. \ref{eq:22}), and then converted it into quantum operators using 
-the Ising Hamiltonian, as proposed by \cite{Tilly2022}.  Prices, returns, and covariance matrix of the assets were calculated using the historical data (see Fig. \ref{fig:data-set}). 
+the Ising Hamiltonian, as proposed by \cite{Tilly2022}.  Prices, returns, and covariance matrix of the assets were calculated using the historical data (see Table \ref{tab:stock_prices}). 
 \begin{equation}
@@
 \centering
-\includegraphics[width=.90\textwidth]{results-vqe.png}
-  \caption{\label{fig:results} Results of the VQE implementation by Buonaiuto et al. \cite{Buonaiuto2023}. The optimal asset allocation for the four assets considered is shown. Extracted from 
+\includegraphics[width=.90\textwidth]{vqe-results.pdf}
+  \caption{\label{fig:results} Results of the VQE implementation by Buonaiuto et al.
+\cite{Buonaiuto2023}.
+The blue line represents the trade of curve between the expected return and the volatility.
+Adapted from
   \cite{Buonaiuto2023}}
 \end{figure*}
+
+The results show that the performance of the algorithm is highly dependent on the ansatz selected and the hardware topology, which dictates the physical arrangement of the qubits and thus, the level of efficiency  the state preparation can be performed with.
+
+Furthermore, the authors mention that a bigger quantum volume is needed to perform deeper circuits without  an exponential 
+increase in the error rate.
 
@@
 
+Future research in NISQ devices with larger quantum volumes, lower error rates, and error mitigation protocols would be crucial 
+to further develop these quantum algorithms.
+
 Finally, although the results of the quantum algorithms applied to financial applications are promising, 
